### 2024-04-05 R/eda-code/EDA_unified.qmd
replacing corrupt file

---
title: "Class Project"
author: Andrew Ruiz
date: "2024-03-13"
output: html_document
---

# load needed packages. make sure they are installed.
```{r}
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(readr)
library(forecast)
library(tsibble) # For time series data wrangling
library(feasts) # For decomposing and visualizing time series
library(RColorBrewer)
```

# This file will contain the EDA code for the class project.
### As the project progresses and the models finalized, the EDA QMD will be updated accordingly.

## MIR heatmap
### This heat map illustrates the timing and intensity of mosquito infection rates (MIR) over the weeks of the year.
#### At least one previous study that a non-zero MIR at or before week 31 is common in years when human cases of EEE are reported. 
```{r}
# Load the saved function
create_heatmap <- readRDS(here("data", "processed-data", "rds","create_heatmap_function.rds"))

# Assuming MIR_by_week is loaded and prepared in this QMD as well
# Now use the loaded function to create the heatmap
MIR_by_week_heatmap <- create_heatmap(MIR_by_week)

# Display the heatmap
MIR_by_week_heatmap
```

## This is a scatter plot of the average number of mosquitoes that were captured in light traps and the mosquito infection rate (MIR).
```{r}
# Load the function from the RDS file
MIRMEL_scatter_plot <- readRDS(here("data", "processed-data", "RDS", "create_scatter_plot_function.rds"))

# Assuming combined_data_mir_all is available in your current session
# Generate the plot
scatter_plot <- create_scatter_plot(combined_data_mir_all)

# Display the plot
print(scatter_plot)

```