### 2024-04-05 products/manuscript/Manuscript.qmd

---
title: "Factors Influencing Abundance & Virus Isolation in Mosquitoes"
subtitle: "Which environmental, ecological, and temporal factors impact mosquito populations"
author: Andrew Ruiz
date: "`r Sys.Date()`"
md_number: true
format:
  docx:
    toc: false
    highlight-style: github
    reference_docx: path/to/your/reference/docx/file.docx
editor: 
  markdown: 
    wrap: 72
bibliography: 
  - ../../assets/project_mada.bib
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(dplyr)
library(readr)

here::set_here()

```

::: {#summary .unnumbered}
# Summary/Abstract

*This project will inform the development of my dissertation prospectus
by identifying significant relationships between variables and guide the
methods development.* Eastern equine encephalitis (EEE) is caused by an
Alphavirus transmitted to humans by the bite of an infected mosquito.
Human infections are rare, but serious. About 30-50% of symptomatic
cases lead to death and those who survive are left with life-long
disability. The transmission cycle for EEE includes two distinct cycles:
an amplifying enzootic cycle in which the virus is transmitted between
the *Culiseta melanura* mosquito and birds and an epizootic cycle in
which bridge vector mosquitoes transmit the EEE virus (EEEv) from birds
to humans and other mammals [@goZoonoticEncephalitidesCaused2014].
Because C. melanura feeds almost exclusively on birds, it is not
considered a direct human threat. However, EEEv isolations in *C.
melanura* are an early warning sign that EEEv is circulating in the
ecosystem. Isolations in bridge vector mosquito species indicate
heightened transmission risk to humans.
:::

# Problem statement

Which environmental,ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts and which
models best describe the relationship between these factors?

**This project will lay the foundation for dissertation research that
will begin in June, 2024.**

# Introduction

## Background

Eastern equine encephalitis (EEE) is caused by an Alphavirus transmitted
to humans by the bite of an infected mosquito. Human infections are
rare, but serious [@lindseyEasternEquineEncephalitis2018]. Humans and
other mammals are dead-end hosts for EEE. About 94% of human EEE
infections are asymptomatic. While symptomatic cases are rare, the
consequences can be severe. Mortality rates associated with symptomatic
infections range from 30 to 50% [@morensEasternEquineEncephalitis2019].
Those who survive often experience lifelong disability that includes
debilitating neurological damage
[@massachusettsdepartmentofpublichealthArbovirusSurveillancePlan2024].
On average, there are 11 cases reported in the US annually. Most occur
in Massachusetts and Florida [@vanderkelenRiskIndexModel2014]. Until
recently, nearly all human EEE cases in Massachusetts occurred in two
counties: Bristol and Plymouth.

Since 2000, there have been 45 human cases recorded in Massachusetts,
resulting in 22 deaths
[@massachusettsdepartmentofpublichealthArbovirusSurveillancePlan2024].
There are no human vaccines for EEE, and treatment consists of
palliative care only. Prevention measures include personal behaviors to
avoid mosquito bites and decreasing the mosquito population through
pesticide use and environmental modifications like removing standing
water
[@centersfordiseasecontrolandpreventionPreventionEasternEquine2022].

The Bristol County Mosquito Control Project (BCMCP) coordinates mosquito
surveillance and testing in the county from June to October. BCMCP has
used the same sentinel collection sites and trapping methods for over 40
years. Once trapped, mosquitoes are sorted and counted by species, and
vector species are submitted to the Massachusetts Department of Health’s
(MDPH) State Lab for PCR virus testing. Results are available within 24
hours of submission. When rates of EEE mosquito infections are above a
defined threshold, MDPH notifies the local boards of health and
recommends preventive measures. Recommended measures include outreach
and education to increase personal protective practices . Additionally,
MDPH may recommend that cities and towns in the affected areas cancel
evening outdoor events and discourage outdoor activities when mosquito
vectors are most active (dawn and dusk). When infection rates are high
enough to indicate imminent human transmission, the Commonwealth of
Massachusetts will recommend and fund aerial pesticide applications over
the affected areas.

## Study area

Bristol County, Massachusetts covers an area of 691 square miles with a
population density of 962 people per square mile. It borders Rhode
Island to the west, Plymouth County, MA to the east, and Norfolk County,
MA to the north. Naushon Island and Martha's Vineyard are across
Buzzard's Bay to the Southeast.

Atlantic White Cedar (AWC) swamps span over 4,300 acres within
Massachusetts, primarily concentrated in Plymouth and Bristol counties.
These swamps serve as critical breeding habitats for *Culiseta melanura*
(MEL)
[@commonwealthofmassachusettsdivisionoffisheries-wildlifeNaturalCommunityFact2007],
which is recognized as the primary enzootic vector for Eastern Equine
Encephalitis (EEE). AWC trees have large root systems that form "crypts"
and hold water throughout the year. These crypts serve as a main habitat
for MEL larval growth and provide shelter during winter months.

Between 2003 and 2022 Plymouth and Bristol counties were the epicenters
for EEE transmission in Massachusetts, accounting for 24 out of the
state's 41 human EEE cases.The Bristol County Mosquito Control Project
(BCMCP) conducts mosquito surveillance and abatement activities for the
county and often coordinates activities with Plymouth County Mosquito
Control Project.

## Questions/Hypotheses to be addressed

*This project will lay the foundation for my dissertation research.*
**The goal for this phase of the project is to answer which
environmental, ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts.** From prior
research we know that the mosquito life cyle is influenced by a number
of factors. Temperature plays a major role in the mosquito life cyle
virus transmission dynamics. As temperature increases, the rate of
larval development also increases and the time between when a mosquito
ingests virus in a bloodmeal and when it can transmit it decreases. The
result is that the mosquito population can grow at faster rate and they
will have more time to transmit the virus. However, MEL larvae develop
in well insulated crypts that are less sensitive to ambient air
temperature changes, so the effect of temperature changes on MEL
development speed may be less pronounced than other vector species that
develop in open, standing water, like ditches and containers.
Precipitation increases the amount of standing water and increases the
chances of larval survival during the mosquito season and also during
the winter months for those species that overwinter in the larval stage.
For EEE, the main enzootic vector, MEL, overwinters as larvae.

While these relationships are well known, they are not well described
for EEE, especially in terms of quantifying the role these two factors
play in EEE transmission. Specifically, there's a gap in our
understanding of how these factors influence the number of mosquitoes
caught in traps and the rate of virus infection among those mosquitoes.
This gap is more pronounced when distinguishing the impacts on enzootic
(mosquitoes that primarily feed on birds) versus bridge vector
(mosquitoes that can transmit the virus from birds to humans) mosquito
species. Even less well described is the effect the factors have on
enzootic vs. bridge vector mosquito species.

Moreover, the influence of other variables remains even less explored.
Birds, especially passerine species, serve as amplifying hosts for EEEV
and WNV, potentially affecting the level of virus present in an area.
Despite their importance in the virus transmission cycle, the effect of
fluctuations in bird populations on the prevalence of these viruses
throughout a season has received minimal attention. Finally, given the
migratory nature of many passerine species, it's worth exploring whether
EEEV activity in the warmer regions of the United States, during the
months preceding mosquito season in Massachusetts, could serve as an
early indicator of virus activity in the state. Understanding these
dynamics is crucial for predicting outbreaks and implementing effective
control measures.

While the following questions will not be addressed in this project, it
will hopefully inform research that will answer the following questions
in dissertation research beginning this summer:

1\. Can an AI-driven model that leverages historic mosquito
surveillance, ecological, and environmental data accurately quantify the
risk of human EEEv infection in southeastern Massachusetts?

2\. What is the potential of machine learning algorithms to identify
early warning signals for EEE outbreaks, enabling timely public health
interventions in Massachusetts to prevent human infections?

3\. How can machine learning models use real-time data on mosquito
activity, environmental factors, and weather to accurately predict when
the risk of EEE transmission to humans is low enough to lift evening
outdoor activity restrictions? \*

## Data collection and processing

### Mosquito surveillance data

The mosquito surveillance dataset encompasses counts by species from all
mosquito traps deployed in Bristol County, MA, from 2007 to 2023. The
raw dataset includes over 45,000 records and includes temporal and
spatial variable. The BCMCP, perform mosquito surveillance from June
through October annually.

During this interval, CDC light traps baited with carbon dioxide are
placed at various sentinel sites each week. Designed to attract
mosquitoes in search of blood meals, light traps serve as a reliable
measure of the actively feeding mosquito population, thereby
establishing them as the standard method for EEE surveillance. Some of
the sentinel site have been used for over 50 years.

Additionally, the county employs gravid traps, which are specifically
aimed at attracting and capturing gravid *Culex* mosquitoes looking for
egg-laying sites. While these traps are primarily used for West Nile
Virus (WNV) surveillance, they hold significant value for EEE monitoring
as well. Given the role of *Culex* species as bridge vectors for EEE,
their capture in gravid traps provides critical insight into the risk of
virus transmission.

The traps are installed in the afternoon and retrieved the following
morning. After the traps are collected, the next are stored in a freezer
to preserve and kill the mosquitoes. The mosquitoes are sorted by
species using a dissecting scope. Counts for each species are recorded
in the database and a portion of the vector species are driven to the
Massachusetts State Laboratory in Boston. The lab tests each mosquito
pool (or batch) for WNV and EEEv using PCR. Results are shared with
BCMCP on the same day of testing. Virus isolation results are recorded
in the collection records.

In addition to mosquito counts by species, the dataset includes the town
or city in which they were collected, the collection date, and whether
the mosquitoes were submitted for virus testing.

### Minimun infection rate (MIR)

One indicator of the level of virus activity present in the mosquito
population can be calculated from the information contained in dataset.
The minimum infection rate (MIR) can be calculated at various temporal
and spatial scales from trap site over one day to all sites over an
entire season. MIR is calculated as: $$
MIR = \left( \frac{\text{Number of positive mosquito pools}}{\text{Total number of mosquitoes tested}} \right) \times 1000
$$ MIR provides an estimate of the prevalence of a specific pathogen
within a mosquito population. While it does not directly measure the
risk to humans, a higher MIR in mosquitoes that commonly bite humans can
indicate a higher risk of transmission. A study pubulihsed in 2007
indicated that in years with human cases in Massachusetts between 1979
and 2004, MIR was ≥1.0 by week 31 [@hachiyaHumanEasternEquine2007]. MIR
allows for the comparison of infection rates across different
geographical areas or time periods. This is crucial for identifying
areas or times of higher risk and for evaluating the effectiveness of
control measures.

![Minimum infection rate by
week](images/MIR_by_week_heatmap.png){fig-align="center"}

The surveillance dataset is structured to conform with state and federal
reporting requirements. As a result, the format of the data must be
transformed to in order to make it ready for analysis. Once the data
structure is corrected, it can be joined to other relevant datasets
using spatial or temporal attributes.

### Weather data

*weather data is currently being reprocessed because I have decided to
switch from daily summaries to hourly summaries. new analysis will be
added in version 5* Weather data played a significant role in this phase
of the analysis. Hourly weather summaries were downloaded from NOAA's
Climate Data Portal. The dataset covers the years 2007-2023 at the
Taunton Airport in Bristol County.

The hourly summaries were used to calculate weather variables, such as
the daily maximum temperature and precipitation. While daily summaries
provide this information will less processing, the hourly summaries
provide more granular data that can be used to calculate additional
variables, such as the number of hours above a certain temperature
threshold or degree-days. Degree-days is a measure of the amount of heat
accumulated over a period of time. It is calculated by summing the
difference between the average daily temperature and a base temperature.
Degree-days are used to estimate the rate of development of insects and
other organisms that are sensitive to temperature. For mosquitoes,
degree-days can be used to estimate the rate of larval development and
the time between when a mosquito ingests a virus and when it can
transmit it [@guStatisticalEstimationDegree2006].

In addition to the weather summaries, drought data will be included in
the models. Drought data was obtained from the US Drought Monitor and
the Federal Emergency Management Agency (FEMA). The data includes weekly
drought severity ratings for the entire United States. The data is
available at the county level and will be used to calculate the number
of weeks in drought conditions and the severity of the drought. For this
project, drought data for Bristol County, MA, will be used and
incorporated with regional drought indicators.

### Human and animal case data

While the goal of this project is to set a foundation for models that
predict human cases of EEE later this year. Human case data will not be
an outcome in the models for this project. However, the case data will
incorporated into the models. Some have hypothesized that EEE activity
in the southern United States could serve as an early warning signal for
EEE activity in Massachusetts. The rationale is that birds that migrate
from the south to Massachusetts could bring the virus with them. In
active EEE years, more infected birds may arrive in Massachusetts and
spread the virus to local mosquitoes at a higher rate. Florida is a
known hotspot for EEE activity and is the only state that reports more
human cases than Massachusetts. The Florida Department of Health reports
EEE activity in animals and humans annually. In addition to the wild and
domestic animal case reports, many mosquito control programs in Florida
maintain flocks of sentinel chickens that are tested for EEE and WNV.
The sentinel chicken data is used to monitor virus activity in the
mosquito population.

### Bird observation data

The North American Breeding Bird Survey (BBS) is conducted annually in
late spring or early summer. The United State Geological Survey (USGS)
maintains the BBS database. The BBS is a long-term, volunteer-driven
bird monitoring program that collects data on bird populations across
North America. The data is used to track the status and trends of bird
populations over time. The BBS data includes information on the number
of birds observed, the location of the observation, and the date of the
observation. The data is collected by trained volunteers who conduct
roadside surveys along predetermined routes. The data is used to
estimate the population size and distribution of bird species across the
continent. The BBS data will be used to estimate the population size and
distribution of bird species in Massachusetts.

Since birds are the amplifying hosts for EEE, their population size and
distribution can have a significant impact on the prevalence of the
virus in the environment. Additionally, the migratory patterns of birds
can influence the spread of the virus across regions. By incorporating
bird observation data into the models, we can better understand the
relationship between bird populations and EEE activity in Massachusetts.

### Temporal scale

For most models in this project, the outcome, MIR, will be calculated by
week. While the main temporal scale for this project is the week, the
data will be aggregated to other temporal scales. The week is the most
common temporal scale for mosquito surveillance data. The week number
will be calculated according to ISO 8601 standards, and stored in a
YYYY-WW format. For example, week 8 in 2019 will be displayed as
2019-08. The week number will provide a common field by which all
datasets with a temporal attribute can be joined. It will also allow for
grouping records and making calculations by week number. This format
helps to standardize temporal scales and make comparisons between years
easier.

While mosquito surveillance and weather data can be aggregated to the
week, other datasets, such as the Annual BBS cannot appropriately be
aggregated by week. Additionally, mosquito surveillance in Massachusetts
is conducted from June to October and EEE activity in Florida occurs
year-round. For these predictors, the temporal scale will be the year.

## Data aquisition

With the exception of the mosquito collection data, all other datasets
are publicly available for use without restrictions.

## Data import and cleaning

Datasets will be loaded into R in a Quarto document to document and the
processing methods will be documented.

The Massachusetts Department of Public Health requires that all mosquito
control districts submit their collection data in a specific format.
This format divides each trap event into separate rows based on mosquito
species and whether they were submitted for virus testing. However, this
format is not useful for certain calculations where a zero count for a
species is not explicitly recorded. For example, even when sorted by
species to calculate the average count per trap, the average could be
inflated since the records will not include trap events with a zero
count of that species. In order to correct this, the data was pivoted to
create a column for every unique "species code" then the "pool size" for
each species code is summed for each trap event. A trap event is when
"town", "date of collection", "trap type" are all the same. For species
codes not included in the orginal table for a trap event, the field will
be assigned a zero value. The resulting table will have one row for
every trap event.

# Methods

*This section is under development and will be completed after analysis
is complete. The following is a general outline of the methods.*

Symptomatic EEE, and to a lesser extent WNV, human cases are rare. On
average, the United States records only about 9 cases of EEE annually.
Notably, Massachusetts reports more cases than any other state except
Florida, highlighting regional disparities in disease occurrence.
However, the incidence rate for neuroinvasive EEE cases in Bristol
County, Massachusetts, is estimated to be only around 0.06 per 100,000
people
[@centersfordiseasecontrolandpreventionPreventionEasternEquine2022]. The
low incidence of human cases poses significant challenges for
traditional statistical models, making it difficult to accurately
estimate human transmission risk. This difficulty is largely due to the
sparse nature of the data and the variability in transmission patterns,
which complicates efforts to predict outbreaks and necessitates
alternative modeling approaches.

As stated earlier, MIR will be the main outcome of interest for this
project.

### Schematic of workflow

*Schematic will be refined in later steps. The doc will also be
reformatted to move the schematic to an appendix* *The image should
appear upon rendering*

```{mermaid}
flowchart TB
    id1["Questions"] --> id2["Data Collection"]
    id2 --> id3["Data Loading and Cleaning"]
    id3 --> id4["Data Cleaning"]
    id4 --> id5["Data Transformation"]
    id5 --> id6["Data Exploration & Visualization"]
    id6 --> id7["Data Modeling"]
    id7 --> id8["Evaluation"]

    id8 -.-> id1
    id8 --> id9[Interpretation]
    id9 --> id10[Communication]

    classDef mainProcess fill:#fff,stroke:#333,stroke-width:2px;

    class id1,id2,id3,id4,id5,id6,id7,id8,id9,id10 mainProcess
```

## Exploratory/Descriptive analysis

It is important to know where the mosquitoes infected with EEE are
captured. Below is a table showing the virus isolations for EEE in
trapped mosquitoes by county between 2014 and 2020.

![](images/eee_isolation_county-04.png){fig-align="center" width="192"}

One factor to consider is if early virus isolations are associated with
cases in humans and other mammals. The table below shows the date of the
first EEE virus isolation in mosquitoes and the number of human of other
animal cases by year.

![](images/eee_isolation_cases-04.png){fig-align="center" width="251"}

Unfortunately, the time horizon is limited by data availability.
Additionally, EEE human infections are rare, so traditional statistical
methods may be inadequate to suggest any relationships. What is clear is
that 2019 was an exceptional year for EEE.

*Culiseta melanura* (MEL) plays an important role in EEE virus
amplification in the enzootic cycle. Understanding the population
dynamics might provide more insight into what is driving the 2019 spike.

The graph below shows the proportion of MEL among all mosquito vector
species trapped in Bristol County, MA.

![](images/mel_proportion.png){fig-align="center"}

From this we see a large proportion of MEL in the traps for 2019. In
fact, they comprise more than half of all vector species for that year.
Another interesting aspect is the high proportion in 2018. MEL
overwinters as larvae and as long as there is sufficient standing water
to last the winter, MEL hatched in the previous fall would survive to
emerge the next year. This suggests it might be important to look at
conditions in the previous year or season for a better understanding.

Since water plays a significant role in the mosquito life cycle, let's
look at the annual precipitation.

![](images/annual_precip-02.png){fig-align="center"}

The graph indicates that the annual precipitation total for 2018 is
higher than the other years. This suggest that there was enough water to
support large MEL larval populations though out the winter.

Another view plots the average annual MEL catch against total
precipitation.

![](images/combined_trends_mel_precip.png){fig-align="center"}

## Basic statistical analysis

The initial tests have been basic regression models.

The first models explored the relationship between virus isolation in
mosquitoes and weather conditions (average temperature by week number
and total precipitation by week). The outcome was whether there was at
least 1 positive pool in a week. The predictor in the first model was
the average maximum daily temperature the week before the pool week. The
logistic model suggest that there was a relationship between the two.
(The R code for the model can be found in the EDA-Code file:
logistic\_.pos_temp_precip.R) The average temperature in the week before
is positively associated with the likelihood of a positive result, with
an odds ratio of approximately 1.069 for each unit increase in
temperature (Estimate = 0.06695, SE = 0.02033, p \< 0.001), indicating
that higher temperatures significantly increase the probability of a
positive outcome. The model's improvement is highlighted by a decrease
in residual deviance to 405.79 from the null deviance of 417.24, upon
incorporating Lagged_Avg_TMAX as a predictor, demonstrating a better fit
to the observed data. However, with a residual deviance of 405.79 on 299
degrees of freedom, it is evident that not all variability in the
outcome is explained, suggesting the presence of other influential
factors. The model's AIC of 409.79 indicates a balance between model
complexity and predictive power, supporting its relative utility in
understanding the data. However, this relationship should be examined
more closely as temperature is likely auto correlated with progressing
weeks during summer months. While precipitation has been demonstrated to
influence the population dynamics it did not show to have any effect on
a positive result.

Another model assessed whether the temperature precipitation in April,
May, and June of each year effected the date of the first positive pool
test. The date of the first positive pool was identified and assigned a
day number from January 1st of that year. The predictors in this linear
model were the average maximum temperature for April, May, and June and
the sum of precipitation for the same time period. None of these
predictors were significant.

There was one other significant result. When MIR is set as the
dependent, continuous variable in a linear regression and the average
maximum temperature for the previous week is used as the predictor, the model suggests there
is a significant inverse relationship. As average temperature increases,
the MIR decreases (model_mir_temp.R). This relationship is
counterintutive, especially given the result of the logistic regression
and requires more examination.

More models assessing the effects of weather and other factors on
positive pools by mosquito species and MIR values will be developed in
Part 4 of the assignment. Future models will examine the the effects of
weather conditions on mosquito population dynamics.

## Statistical analysis

*UPDATE for Pt 4* *a better source for weather data was found and it
being processed.* *the analysis will be updated in the next version of
the document*

![](images/correlation_plot.png){fig-align="center"}

# Results

*updated in later phases of the project*

## Full analysis

*These will be generated in Pt 5* 
# Discussion 
*this section will be
completed when analysis is done*

## Strengths and Limitations

*this section will be updated after analysis is complete* The accuracy
and reliability of the mosquito collection data is excellent. The same
procedures and sentinel sites have were in place for the entire
analytical horizon. Further, there were only two people responsible for
sorting, identifying and counting species over the entire period.

MIR is not a perfect measure of virus activity in mosquitoes. It tends
to underestimate the level of virus circulating in the environment.
However, other measures require equal pool numbers for testing. There is
a correction factor for variable pool numbers and I will explore this
option in later phases of this project.

The spatial scale is an issue with the human and veterinary cases data.
Human case data is released only on a county scale to ensure patient
privacy. Some veterinary case data is available on the town-level, but
most remains only on the county-level. These limitations are barriers to
accurate spatial analysis. The project will rely on virus isolation in
mosquitoes to model risk. The available bird data relies heavily on
crowd-sourced reports. There is one standardized survey (Breeding Bird
Survey) that seems to be more accurate and reliable. However, this data
is not available for every year.

## Conclusions *this section will be completed when analysis is done*

# References