### 2024-04-21 R/analysis-code/analysis.qmd
Part 5
Analysis is complete. Readme files and removal of outdated datasets and codes will be removed throughout the week. Additionally, the manuscript will be finalized this weel

---
title: "statistical-analysis"
author: "Andrew Ruiz"
format: html
editor: source
---

# Load the necessary libraries for the analysis

```{r}
library(here)
library(dplyr)
library(lubridate)
library(readr)
library(stargazer)
library(mgcv)
library(tidyr)
library(caret)
library(randomForest)
library(tidymodels)
library(ranger)
library(dials)
library(vip)
library(ggplot2)
library(boot)
library(yardstick)
library(glmnet) # For LASSO
library(dials)
library(mlr3)
library(data.table)
library(Matrix)
library(gbm)
```

## First logistic model

#### Effect of temperature and precipitation the week prior to mosquito collection on the likelihood of observing at least one positive mosquito test result in a week

```{r}
# Load the PCR data set which includes information on mosquito pools tested for viruses, 
# results of those tests, and the date of collection.
pcr <- read_csv(here("data", "processed-data", "pcr_data_expanded_filtered.csv"))

# Load the weather data set which includes daily maximum temperatures (TMAX) and other weather variables.
wx_selected <- read_csv(here("data", "processed-data", "wx_selected.csv"))

# Aggregate the weather data by week to calculate the average maximum temperature for each week.
# This simplifies the daily temperature data into a more manageable weekly summary,
# making it easier to correlate with weekly mosquito test results.
wx_selected <- wx_selected %>%
  group_by(week_num) %>%
  summarise(
    Avg_TMAX = mean(TMAX, na.rm = TRUE),
    Total_PRCP = sum(PRCP, na.rm = TRUE)
  ) %>%
  ungroup()

# View the first few rows to confirm it worked
head(wx_selected)

# Create a lagged version of the average weekly maximum temperature to explore the potential delayed effect of temperature on mosquito test results.
# The lag function shifts the temperature data by one week, assuming that the impact of temperature on virus transmission might not be immediate.
weather_weekly <- wx_selected %>%
  arrange(week_num) %>%
  mutate(Lagged_Avg_TMAX = lag(Avg_TMAX, 1))

weather_weekly <- wx_selected %>%
  arrange(week_num) %>%
  mutate(
    Lagged_Avg_TMAX = lag(Avg_TMAX, 1),
    Lagged_Total_PRCP = lag(Total_PRCP, 1)
  )

weather_weekly <- wx_selected %>%
  arrange(week_num) %>%
  mutate(
    Lagged_Avg_TMAX = lag(Avg_TMAX, 1),
    Lagged_Total_PRCP = lag(Total_PRCP, 1),
    Binary_PRCP = if_else(Lagged_Total_PRCP > 0, 1, 0)
  )
# Aggregate the PCR data by week to determine if there was at least one positive result in each week.
# This transformation reduces the detailed test results into a weekly summary,
# indicating the presence or absence of positive tests each week.
pcr_weekly <- pcr %>%
  group_by(week_num) %>%
  summarise(Positive_Result = max(as.integer(Is_Positive), na.rm = TRUE))

# Merge the aggregated PCR data with the lagged weather data by matching their week numbers.
# This combined data set enables the analysis of the relationship between weekly weather conditions and mosquito test results.
combined_wx_pcr <- merge(pcr_weekly, weather_weekly, by = "week_num")

# Fit a logistic regression model to examine the influence of lagged average weekly maximum temperature on the likelihood of observing at least one positive mosquito test result in a week.
# The binomial family specifies that the dependent variable (Positive_Result) is binary,
# making logistic regression an appropriate modeling choice for this binary outcome.
logistic_model_temp <- glm(Positive_Result ~ Lagged_Avg_TMAX + Binary_PRCP, family = binomial, data = combined_wx_pcr)

# Output the summary of the logistic regression model to review the estimated coefficients, their statistical significance, and the overall model fit.
# This summary provides insights into the relationship between temperature and the probability of positive mosquito test results.
summary(logistic_model_temp)

# Output a LaTeX-formatted table to the console
stargazer(logistic_model_temp, type = "text")

par(mfrow = c(2, 2))
plot(logistic_model_temp)

```

```{r}
pcr <- read_csv(here("data", "processed-data", "pcr_data_expanded_filtered.csv"))
wx_selected <- read_csv(here("data", "processed-data", "wx_selected.csv"))

# Step 2: Calculate the average TMAX for April, May, and June separately
avg_temp_by_month <- wx_selected %>%
  filter(month(DATE) %in% c(4, 5, 6)) %>%
  group_by(Year = year(DATE), Month = month(DATE)) %>%
  summarise(Avg_TMAX = mean(TMAX, na.rm = TRUE), .groups = 'drop')

# Pivot the average temperatures into separate columns for each month
avg_temp_by_month_wide <- avg_temp_by_month %>%
  pivot_wider(names_from = Month, values_from = Avg_TMAX, names_prefix = "Avg_TMAX_M") %>%
  # Ensure all years are present even if some data might be missing
  complete(Year, fill = list(Avg_TMAX_M4 = NA, Avg_TMAX_M5 = NA, Avg_TMAX_M6 = NA))

# Step 4: Identify the first positive test date each year in mosquito_data
first_positive_dates <- pcr %>%
  filter(Result == "Positive") %>%
  group_by(Year = year(`Collection Date`)) %>%
  summarise(First_Positive_Date = min(`Collection Date`), .groups = 'drop')

# Step 5: Merge the datasets on Year
combined_data_firstpos <- left_join(first_positive_dates, avg_temp_by_month_wide, by = "Year")

# Step 6: Convert First_Positive_Date to a numerical value (e.g., day of the year)
combined_data_firstpos <- combined_data_firstpos %>%
  mutate(First_Positive_DOY = yday(First_Positive_Date))

# Explore the relationship via linear regression with Avg_TMAX_M4, Avg_TMAX_M5, Avg_TMAX_M6 as predictors
model2 <- lm(First_Positive_DOY ~ Avg_TMAX_M4 + Avg_TMAX_M5 + Avg_TMAX_M6, data = combined_data_firstpos)
summary(model2)

par(mfrow = c(2, 2))
plot(model2)

```

# Does the total precipitation in April, May, and June influence the date of the first positive test?

```{r}
wx_selected <- read_csv(here("data", "processed-data", "wx_selected.csv"))

# Calculate average TMAX and total PRCP for April, May, and June separately
avg_temp_and_prcp <- wx_selected %>%
  filter(month(DATE) %in% c(4, 5, 6)) %>%
  group_by(Year = year(DATE), Month = month(DATE)) %>%
  summarise(
    Avg_TMAX = mean(TMAX, na.rm = TRUE),
    Total_PRCP = sum(PRCP, na.rm = TRUE),
    .groups = 'drop'
  )

# Continue from the summarise step
wide_data <- avg_temp_and_prcp %>%
  # Convert to a long format to distinguish between Avg_TMAX and Total_PRCP
  pivot_longer(cols = c(Avg_TMAX, Total_PRCP), names_to = "Measure", values_to = "Value") %>%
  # Create a combined name for later pivot_wider
  unite("NewName", Measure, Month, sep = "_M") %>%
  # Pivot wider with a single names_prefix
  pivot_wider(names_from = NewName, values_from = Value) %>%
  complete(Year, fill = list(Avg_TMAX_M4 = NA, Avg_TMAX_M5 = NA, Avg_TMAX_M6 = NA, Total_PRCP_M4 = NA, Total_PRCP_M5 = NA, Total_PRCP_M6 = NA))

# Load and prepare mosquito data
first_positive_dates <- pcr_data_expanded_filtered %>%
  filter(Result == "Positive") %>%
  group_by(Year = year(`Collection Date`)) %>%
  summarise(First_Positive_Date = min(`Collection Date`), .groups = 'drop')

# Merge the datasets on Year
combined_data <- left_join(first_positive_dates, wide_data, by = "Year")

# Convert First_Positive_Date to a numerical value (e.g., day of the year)
combined_data <- combined_data %>%
  mutate(First_Positive_DOY = yday(First_Positive_Date))

# Explore the relationship via linear regression
model <- lm(First_Positive_DOY ~ Total_PRCP_M4 + Total_PRCP_M5 + Total_PRCP_M6, data = combined_data)
summary(model)

```

## Two-stage modeling approach using Generalized Additive Models (GAMs) to understand the factors influencing Mosquito Infection Rate (MIR).

### First stage: zero inflation model

#### This model predicts whether MIR is greater than zero using smoothed functions of several predictors, such as the Drought Severity Composite Index for the area (DSCI_ma), average mosquito larvae counts (Average_MEL), and the average current hourly wind speed (Avg_HourlyWindSpeed_current).

##### *The package required for this is mgcv*

```{r}
# load the dataset with a unique dataframe name
mir_2part <- read_csv(here("data", "processed-data", "joined_data_mir_co2.csv"))

#define the model
zero_gam_model <- gam(I(MIR > 0) ~ s(DSCI_ma) + s(Average_MEL) + s(Avg_HourlyWindSpeed_current),
                      family = binomial(link = "logit"), 
                      data = mir_2part)

summary(zero_gam_model)
plot(zero_gam_model)

# Define the null model with only an intercept
null_gam_model <- gam(I(MIR > 0) ~ 1,
                      family = binomial(link = "logit"), 
                      data = mir_2part)

summary(null_gam_model)

# Calculate the AIC for the null model
aic_null_gam_model <- AIC(null_gam_model)
aic_null_gam_model

aic_zero_gam_model <- AIC(zero_gam_model)
aic_zero_gam_model

# # load the dataset with a unique dataframe name
# mir_2part_all <- read_csv(here("data", "processed-data", "all_mir_clean_joined.csv"))

#define the model
zero_gam_model <- gam(I(MIR > 0) ~ s(DSCI_ma) + s(Average_MEL) + s(Avg_HourlyWindSpeed_current),
                      family = binomial(link = "logit"), 
                      data = mir_2part)

summary(zero_gam_model)
plot(zero_gam_model)

# Define the null model with only an intercept
null_gam_model <- gam(I(MIR > 0) ~ 1,
                      family = binomial(link = "logit"), 
                      data = mir_2part)

summary(null_gam_model)

# Calculate the AIC for the null model
aic_null_gam_model <- AIC(null_gam_model)
aic_null_gam_model

aic_zero_gam_model <- AIC(zero_gam_model)
aic_zero_gam_model
```

```{r}

# Response variable
response_var <- "MIR"

# Specified predictors for modeling
specified_predictors <- c("Avg_TMAX", "Total_PRCP", "Hrs_Above_80_lag", "Hrs_Above_75_lag",
                          "Hrs_Below_50_lag", "Hrs_Between_50_86_lag", "Avg_HourlyDryBulbTemp_lag",
                          "Avg_HourlyWetBulbTemp_lag", "Hrs_Below_50_current", "DSCI_fema",
                          "Avg_HourlyRelativeHumidity_current", "Avg_HourlyWetBulbTemperature_current", 
                          "Avg_HourlyWindSpeed_current", "None.x", "DSCI_fema", "average_daylight_hours", "Average_MEL", "MEL_Prop")

# Initiate lists to store models and summaries
gam_models_zero <- list()
gam_models_positive <- list()
summary_models_zero <- list()
summary_models_positive <- list()

# Loop through each specified predictor
for (predictor in specified_predictors) {
  # Check if predictor exists in the dataset
  if (!predictor %in% names(mir_2part)) {
    cat("Skipping", predictor, "because it is not in the dataset.\n")
    next
  }

  # Zero-inflation part
  formula_zero <- as.formula(paste0("I(", response_var, " > 0) ~ s(", predictor, ")"))
  zero_model <- gam(formula_zero, data = mir_2part, family = binomial(link = "logit"))
  gam_models_zero[[predictor]] <- zero_model
  summary_models_zero[[predictor]] <- summary(zero_model)

  cat("Zero-inflation model for", predictor, "\n")

  # Positive counts part
  if (any(mir_2part[[response_var]] > 0)) {
    positive_data <- mir_2part[mir_2part[[response_var]] > 0, ]
    formula_positive <- as.formula(paste0(response_var, " ~ s(", predictor, ")"))
    positive_model <- gam(formula_positive, data = positive_data, family = gaussian())
    gam_models_positive[[predictor]] <- positive_model
    summary_models_positive[[predictor]] <- summary(positive_model)

    cat("Positive model for", predictor, "\n")
  }

  # print or plot the summaries here...
  print(summary(zero_model))
  print(summary(positive_model))
  plot(zero_model)
  plot(positive_model)
}

### ALL MIR###
# Response variable
response_var <- "MIR"

# Specified predictors for modeling
specified_predictors <- c("Avg_TMAX", "Total_PRCP", "Hrs_Above_80_lag", "Hrs_Above_75_lag",
                          "Hrs_Below_50_lag", "Hrs_Between_50_86_lag", "Avg_HourlyDryBulbTemp_lag",
                          "Avg_HourlyWetBulbTemp_lag", "Hrs_Below_50_current", "DSCI_fema",
                          "Avg_HourlyRelativeHumidity_current", "Avg_HourlyWetBulbTemperature_current", 
                          "Avg_HourlyWindSpeed_current", "None.x", "DSCI_fema", "average_daylight_hours", "Average_MEL", "MEL_Prop")

# Initiate lists to store models and summaries
gam_models_zero <- list()
gam_models_positive <- list()
summary_models_zero <- list()
summary_models_positive <- list()

# Loop through each specified predictor
for (predictor in specified_predictors) {
  # Check if predictor exists in the dataset
  if (!predictor %in% names(mir_2part_all)) {
    cat("Skipping", predictor, "because it is not in the dataset.\n")
    next
  }

  # Zero-inflation part
  formula_zero <- as.formula(paste0("I(", response_var, " > 0) ~ s(", predictor, ")"))
  zero_model <- gam(formula_zero, data = mir_2part_all, family = binomial(link = "logit"))
  gam_models_zero[[predictor]] <- zero_model
  summary_models_zero[[predictor]] <- summary(zero_model)

  cat("Zero-inflation model for", predictor, "\n")

  # Positive counts part
  if (any(mir_2part_all[[response_var]] > 0)) {
    positive_data <- mir_2part_all[mir_2part_all[[response_var]] > 0, ]
    formula_positive <- as.formula(paste0(response_var, " ~ s(", predictor, ")"))
    positive_model <- gam(formula_positive, data = positive_data, family = gaussian())
    gam_models_positive[[predictor]] <- positive_model
    summary_models_positive[[predictor]] <- summary(positive_model)

    cat("Positive model for", predictor, "\n")
  }

  # print or plot the summaries here...
  print(summary(zero_model))
  print(summary(positive_model))
  plot(zero_model)
  plot(positive_model)
}
```

#### Based on the results above, the following have been identified as potential predictors of MIR:

##### This shows the model and the p-value
###### MIR ~ s(Hrs_Above_80_lag) 0.0398
###### MIR ~ s(Hrs_Above_75_lag) 0.0213
###### MIR ~ s(Hrs_Below_50_lag) 0.00489
###### MIR ~ s(Avg_HourlyDryBulbTemp_lag) 0.0301
###### MIR ~ s(Avg_HourlyWetBulbTemp_lag) 0.00928
###### MIR ~ s(Hrs_Below_50_current) 0.000593
###### MIR ~ s(Avg_HourlyWetBulbTemperature_current) 0.028
###### MIR ~ s(Avg_HourlyWindSpeed_current) 0.0369
###### MIR ~ s(MEL_Prop) 0.00867

###### I(MIR > 0) ~ s(Avg_HourlyWindSpeed_current) 0.00483
###### I(MIR > 0) ~ s(None.x) 0.000935
###### I(MIR > 0) ~ s(average_daylight_hours) 1.15e-06
######I(MIR > 0) ~ s(Average_MEL) 4.75e-05
######I(MIR > 0) ~ s(MEL_Prop) 5.11e-05
###### I(MIR > 0) ~ s(DSCI_ma) 0.0174

### There is some concern that dividing the data into two parts (observations >0 for part 2) may reduce the power of the analysis.

### Now use a GAM with all records, where MIR >= 0
```{r}
# Response variable
response_var <- "MIR"

# Specified predictors for modeling
specified_predictors <- c("Avg_TMAX", "Total_PRCP", "Hrs_Above_80_lag", "Hrs_Above_75_lag",
                          "Hrs_Below_50_lag", "Hrs_Between_50_86_lag", "Avg_HourlyDryBulbTemp_lag",
                          "Avg_HourlyWetBulbTemp_lag", "Hrs_Below_50_current", "DSCI_fema",
                          "Avg_HourlyRelativeHumidity_current", "Avg_HourlyWetBulbTemperature_current", 
                          "Avg_HourlyWindSpeed_current", "None.x", "None.y",  "DSCI_ma", "average_daylight_hours", 
                          "Average_MEL", "MEL_Prop")

# Initiate lists to store models and summaries
gam_models <- list()
summary_models <- list()

# Loop through each specified predictor
for (predictor in specified_predictors) {
  # Check if predictor exists in the dataset
  if (!predictor %in% names(mir_2part_all)) {
    cat("Skipping", predictor, "because it is not in the dataset.\n")
    next
  }

  # Model with all records
  formula <- as.formula(paste0(response_var, " ~ s(", predictor, ")"))
  model <- gam(formula, data = mir_2part_all, family = gaussian())
  gam_models[[predictor]] <- model
  summary_models[[predictor]] <- summary(model)

  cat("Model for", predictor, "\n")

  # Print or plot the summaries here...
  print(summary(model))
  plot(model)
}
```

### The all records model shows the following predictors as significant:
#### MIR ~ s(Hrs_Below_50_lag) 0.0105
#### MIR ~ s(Hrs_Below_50_current) 0.00333
#### MIR ~ s(DSCI_fema) 0.0214
#### MIR ~ s(Avg_HourlyRelativeHumidity_current) 0.00936
#### MIR ~ s(Avg_HourlyWindSpeed_current) 0.00366
#### MIR ~ s(average_daylight_hours) <2e-16
#### MIR ~ s(Average_MEL) 0.0258
#### MIR ~ s(MEL_Prop) <2e-16
## Give the concern with splitting the data further, we will continue with the all records model.

### Now that we have a better idea of the relationships and potential significance.
#### We will have another look at the correlation matrix for the predictors.

```{r}
library(reshape2) # or library(data.table)
library(plotly)

# Assuming continuous_vars is a dataframe with just the continuous variables you want to correlate
# Create a correlation matrix
cor_matrix <- cor(mir_2part %>% 
                    select(Hrs_Above_80_lag, Hrs_Above_75_lag, Hrs_Below_50_lag, 
                           Avg_HourlyDryBulbTemp_lag, Avg_HourlyWetBulbTemp_lag, 
                           Hrs_Below_50_current, Avg_HourlyWetBulbTemperature_current, 
                           Avg_HourlyWindSpeed_current, None.x, average_daylight_hours, 
                           Average_MEL, MEL_Prop, DSCI_ma),
                  use = "complete.obs")  # 'use' parameter handles missing values

# Melt the correlation matrix into a long format
cor_data <- melt(cor_matrix, na.rm = TRUE)

# Ensure cor_data has the expected column names after melting
if (!all(c("Var1", "Var2") %in% names(cor_data))) {
  stop("Var1 and Var2 not found in melted data. Check the melting process.")
}

# Create the interactive heatmap using plotly
heatmap_plot <- plot_ly(
  data = cor_data,
  x = ~Var1,
  y = ~Var2,
  z = ~value,
  type = 'heatmap',
  colorscale = 'Portland'
) %>%
  layout(
    title = 'Correlation Matrix',
    xaxis = list(title = 'Variable 1', tickangle = 45),
    yaxis = list(title = 'Variable 2', autorange = "reversed")
  )

# If you're running this in an interactive R session, this will display the plot
heatmap_plot
```
#### While correlation between predictors is a concern for the GAM model, it is not as much of a concern for the random forest. 
#### I will create a new variable that represents the difference between the average dry bulb temperature and the average wet bulb temperature.
```{r}
# Create a new dataframe by copying mir_2part
mir_2part_updated <- mir_2part

# Add the new variable to the new dataframe
mir_2part_updated$Dry_Wet_Diff <- mir_2part_updated$Avg_HourlyDryBulbTemp_lag - mir_2part_updated$Avg_HourlyWetBulbTemp_lag

# Now, mir_2part remains unchanged, and mir_2part_updated includes the new variable.
```

## 1 part GAM
```{r}
# Load necessary libraries
library(tidymodels)
library(dplyr)
library(mgcv) # For GAM
library(readr) # For read_csv

# Read the dataset
mir_1part_all <- read_csv(here("data", "processed-data", "all_mir_clean_joined.csv"))

# Define predictors for the continuous outcome
predictors_part2 <- c("Hrs_Below_50_lag", 
                      "Hrs_Below_50_current", "Avg_HourlyWetBulbTemperature_current", 
                      "Avg_HourlyWindSpeed_current", "MEL_Prop", "average_daylight_hours", "Average_MEL")

# Prepare the dataset for continuous outcome
mir_1part_all_continuous <- mir_1part_all %>%
  select(MIR, all_of(predictors_part2)) %>%
  na.omit()

# Split the data
set.seed(123)
data_split <- initial_split(mir_1part_all_continuous, prop = 0.75)
train_data_continuous <- training(data_split)
test_data_gam1 <- testing(data_split)

# Calculate the mean of the response variable for the entire dataset
mean_response <- mean(mir_1part_all_continuous$MIR)

# Create predictions using the mean response for the entire dataset
null_predictions <- rep(mean_response, nrow(mir_1part_all_continuous))

# Calculate RMSE for the null model
null_rmse <- sqrt(mean((null_predictions - mir_1part_all_continuous$MIR)^2))
cat("Null Model RMSE (Entire Dataset):", null_rmse, "\n")

# Training for Continuous Outcome
formula_gam1 <- as.formula(paste("MIR ~", paste("s(", predictors_part2, ")", collapse = " + ")))
model_gam1 <- gam(formula_gam1, data = train_data_continuous, family = gaussian())

# Evaluate on Training Set (Continuous)
predictions_train_gam1 <- predict(model_gam1, newdata = train_data_continuous)
rmse_train_gam1 <- sqrt(mean((predictions_train_gam1 - train_data_continuous$MIR)^2))
cat("Training Set RMSE (Continuous Outcome):", rmse_train_gam1, "\n")

# Evaluate on Test Set (Continuous Outcome)
predictions_test_gam1 <- predict(model_gam1, newdata = test_data_gam1)
rmse_test_gam1 <- sqrt(mean((predictions_test_gam1 - test_data_gam1$MIR)^2))
cat("Test Set RMSE (Continuous Outcome):", rmse_test_gam1, "\n")

# Visualize model effects for insights
plot(model_gam1)

### Boot strapping
# Define the bootstrapping statistic function
boot_statistic <- function(data, indices) {
  # Extract the bootstrap sample
  boot_sample <- data[indices, ]

  # Fit the GAM model on the bootstrap sample
  gam_boot <- gam(formula_gam1, data = boot_sample, family = gaussian())

  # Predict on the out-of-bag (OOB) data
  oob_data <- data[-indices, ]
  predictions_oob <- predict(gam_boot, newdata = oob_data)

  # Calculate RMSE on the OOB predictions
  sqrt(mean((predictions_oob - oob_data$MIR)^2))
}

# Perform bootstrapping
set.seed(123)  # Ensure reproducibility
bootstrap_results <- boot(
  data = train_data_continuous,
  statistic = boot_statistic,
  R = 1000  # For example, perform 1000 bootstrap replicates
)

# Extract the bootstrapped RMSE and calculate the mean and standard deviation
bootstrapped_rmse <- bootstrap_results$t
mean_bootstrapped_rmse <- mean(bootstrapped_rmse)
sd_bootstrapped_rmse <- sd(bootstrapped_rmse)

# Display bootstrapped RMSE results
cat("Bootstrapped RMSE: Mean=", mean_bootstrapped_rmse, ", SD=", sd_bootstrapped_rmse, "\n")

# # Calculate confidence intervals for the bootstrapped RMSE
# boot_ci <- boot.ci(bootstrap_results, type = "basic")
# cat("95% CI for RMSE: Lower =", boot_ci$basic[4], ", Upper =", boot_ci$basic[5], "\n")

# Calculate percentile confidence intervals for the bootstrapped RMSE
boot_ci_percentile <- boot.ci(bootstrap_results, type = "perc")

# Display the percentile confidence intervals for the bootstrapped RMSE
cat("95% CI for RMSE (Percentile): Lower =", boot_ci_percentile$percent[4], ", Upper =", boot_ci_percentile$percent[5], "\n")
```

### Residual plots for continuous outcome:
```{r}
# Step 1: Calculate residuals
train_residuals <- residuals(model_gam1, type = "response")

# Step 2: Create residual plot
plot(predictions_train_gam1, train_residuals,
     xlab = "Predicted Values",
     ylab = "Residuals",
     main = "Residual Plot (Training Set)")

# Step 3: Create predicted vs. actual plot for training set
plot(predictions_train_gam1, train_data_continuous$MIR,
     xlab = "Predicted Values",
     ylab = "Actual Values",
     main = "Predicted vs. Actual Plot (Training Set)")

plot(predictions_test_gam1, test_data_gam1$MIR,
     xlab = "Predicted Values",
     ylab = "Actual Values",
     main = "Predicted vs. Actual Plot (Test Set)")

```

## 1 Part Random Forest
```{r}
# Load the dataset
mir_1part_all <- read_csv(here("data", "processed-data", "all_mir_clean_joined.csv"))

# Define predictors for the continuous outcome
predictors <- c("Hrs_Below_50_lag", "Hrs_Below_50_current", 
                "Avg_HourlyWetBulbTemperature_current", "Avg_HourlyWindSpeed_current", 
                "MEL_Prop", "average_daylight_hours", "Average_MEL")

# Prepare the dataset for continuous outcome
train_data_continuous <- mir_1part_all %>%
  select(MIR, all_of(predictors)) %>%
  na.omit()

# Split the data
set.seed(123)
data_split <- initial_split(train_data_continuous, prop = 0.75)
train_data_continuous <- training(data_split)
test_data_continuous <- testing(data_split)

# Define the tuning grid
tuning_grid <- grid_regular(
  mtry(range = c(1, length(predictors))),
  min_n(range = c(5, 25)), # Adjusted based on preliminary insights
  levels = 10 # Increased resolution
)

# tuning_grid <- grid_regular(
#   mtry(range = c(1, length(predictors))),
#   min_n(range = c(1, 21)),
#   levels = 7
# )

# Define cross-validation folds
cv_folds_continuous <- vfold_cv(train_data_continuous, v = 5)

# Define the Random Forest model specification
rf_model_spec_continuous <- rand_forest(trees = 500, mtry = tune(), min_n = tune()) %>%
  set_engine("ranger", importance = "impurity", seed = 123) %>%
  set_mode("regression")

# Create the recipe
rf_recipe_continuous <- recipe(MIR ~ ., data = train_data_continuous)

# Workflow setup
rf_workflow_continuous <- workflow() %>%
  add_recipe(rf_recipe_continuous) %>%
  add_model(rf_model_spec_continuous)

# Tune the model
tuning_results_continuous <- tune_grid(
  rf_workflow_continuous,
  resamples = cv_folds_continuous,
  grid = tuning_grid,
  metrics = metric_set(yardstick::rmse)
)

# Extract the best hyperparameters
best_hps_continuous <- select_best(tuning_results_continuous, metric = "rmse")

# Fit the final model with the best hyperparameters on the full training dataset
final_fit_continuous <- fit(
  rf_workflow_continuous %>% finalize_workflow(best_hps_continuous),
  data = train_data_continuous
)

### Calculate RMSE for the training and testing sets

# Predict on the training and testing set
train_preds <- predict(final_fit_continuous, new_data = train_data_continuous)$.pred
test_preds <- predict(final_fit_continuous, new_data = test_data_continuous)$.pred

# You need to create tibbles that combine the actual values with the predictions
train_results <- tibble(
  truth = train_data_continuous$MIR,
  estimate = train_preds
)

test_results <- tibble(
  truth = test_data_continuous$MIR,
  estimate = test_preds
)

# Now, you can correctly calculate RMSE for both the training and testing sets
train_rmse_result_rf <- rmse(train_results, truth = truth, estimate = estimate)
test_rmse_result_rf <- rmse(test_results, truth = truth, estimate = estimate)

# Display the RMSE results
cat("Training Set RMSE:", train_rmse_result_rf$.estimate, "\n")
cat("Testing Set RMSE:", test_rmse_result_rf$.estimate, "\n")

# Continue with bootstrapping to evaluate model's robustness
set.seed(123) # for reproducibility
bootstrap_results <- boot(
  data = train_data_continuous,
  statistic = boot_fit_and_evaluate,
  R = 100 # number of bootstrap resamples
)

# Calculate the bootstrapped RMSE
mean_rmse_boot_rf <- mean(bootstrap_results$t)
sd_rmse_boot_rf <- sd(bootstrap_results$t)
ci_rmse_boot_rf <- boot.ci(bootstrap_results, type = "perc")  # Percentile interval

# Extracting the correct CI values for percentile CI
lower_bound <- ci_rmse_boot_rf$percent[4]
upper_bound <- ci_rmse_boot_rf$percent[5]

# Output the bootstrapping results
cat("Bootstrapped RMSE: Mean=", mean_rmse_boot_rf, ", SD=", sd_rmse_boot_rf, "\n")
cat("95% CI for RMSE (Percentile): ", lower_bound, "-", upper_bound, "\n")

autoplot(tuning_results_continuous)
```
### The random forest model has evidence of overfitting as the RMSE on the test set is higher than the training set.

## LASSO is not appropriate for non-linear relationships.

### The LASSO model suggests overfitting as well. The RMSE on the test set is higher than the training set.
### We will try modifying the tuning. 
```{r}
library(tidymodels)
library(readr) # For read_csv
library(yardstick) # For metrics

# Load the dataset
mir_1part_all <- read_csv(here("data", "processed-data", "all_mir_clean_joined.csv"))

# Define predictors for the continuous outcome
predictors <- c("Hrs_Below_50_lag", "Hrs_Below_50_current", 
                "Avg_HourlyWetBulbTemperature_current", "Avg_HourlyWindSpeed_current", 
                "MEL_Prop", "average_daylight_hours", "Average_MEL")

# Prepare the dataset for continuous outcome
train_data_continuous <- mir_1part_all %>%
  select(MIR, all_of(predictors)) %>%
  na.omit()

# Split the data
set.seed(123)
data_split <- initial_split(train_data_continuous, prop = 0.75)
train_data_continuous <- training(data_split)
test_data_continuous <- testing(data_split)

# Define a more refined tuning grid for LASSO with a wider range of penalty values
tuning_grid_lasso <- expand.grid(
  penalty = 10^seq(-10, 1, length.out = 50)  # Increase the range and resolution of the penalty parameter
)

# Increase the number of cross-validation folds for a more robust evaluation
cv_folds_continuous <- vfold_cv(train_data_continuous, v = 10)

# Define the LASSO model specification
lasso_model_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet") %>%
  set_mode("regression")

# Create the recipe
#lasso_recipe <- recipe(MIR ~ ., data = train_data_continuous)
lasso_recipe <- recipe(MIR ~ ., data = train_data_continuous) %>%
  update_role(all_of(predictors), new_role = "predictor") %>%
  step_poly(all_of(predictors), degree = 2)

# Workflow setup
lasso_workflow <- workflow() %>%
  add_recipe(lasso_recipe) %>%
  add_model(lasso_model_spec)

# Tune the model using the refined tuning grid and increased folds
tuning_results_lasso <- tune_grid(
  lasso_workflow,
  resamples = cv_folds_continuous,
  grid = tuning_grid_lasso,
  metrics = metric_set(yardstick::rmse)
)

# Extract the best hyperparameters
best_hps_lasso <- select_best(tuning_results_lasso, metric = "rmse")

# Fit the final model with the best hyperparameters on the full training dataset
final_fit_lasso <- fit(
  lasso_workflow %>% finalize_workflow(best_hps_lasso),
  data = train_data_continuous
)

# Predict on the training and testing set using the final LASSO model
train_preds_lasso <- predict(final_fit_lasso, new_data = train_data_continuous)$.pred
test_preds_lasso <- predict(final_fit_lasso, new_data = test_data_continuous)$.pred

# Calculate RMSE for the training and testing sets using the predictions
train_rmse_result_lasso <- yardstick::rmse(data.frame(truth = train_data_continuous$MIR, estimate = train_preds_lasso), truth = truth, estimate = estimate)
test_rmse_result_lasso <- yardstick::rmse(data.frame(truth = test_data_continuous$MIR, estimate = test_preds_lasso), truth = truth, estimate = estimate)

# Display the RMSE results
cat("Training Set RMSE (LASSO):", train_rmse_result_lasso$.estimate, "\n")
cat("Testing Set RMSE (LASSO):", test_rmse_result_lasso$.estimate, "\n")

# Visualize the tuning results
autoplot(tuning_results_lasso)

```
### There still seems to be overfitting. The RMSE on the test set is higher than the training set.
### We will try a different model

# Generalized Additive Models (GAMs) with sparsity
### GAMs are flexible models that can capture non-linear relationships between predictors and the outcome.
### Acts similar to LASSO adding a penalty similar to LASSO to enforce sparsity and variable selection
```{r}
library(tidymodels)
library(mgcv)
library(readr) # For read_csv

# Load the dataset
mir_1part_all <- read_csv(here("data", "processed-data", "all_mir_clean_joined.csv"))

# Prepare the dataset for continuous outcome
train_data_continuous <- mir_1part_all %>%
  select(MIR, Hrs_Below_50_lag, Hrs_Below_50_current, Avg_HourlyWetBulbTemperature_current, 
         Avg_HourlyWindSpeed_current, MEL_Prop, average_daylight_hours, Average_MEL) %>%
  na.omit()

# Split the data
set.seed(123)
data_split <- initial_split(train_data_continuous, prop = 0.75)
train_data_continuous <- training(data_split)
test_data_continuous <- testing(data_split)

# Define cross-validation folds
cv_folds_continuous <- vfold_cv(train_data_continuous, v = 5)

# Placeholder for storing CV results
cv_results <- vector("list", length(cv_folds_continuous$splits))

# Loop through CV folds
for (i in seq_along(cv_folds_continuous$splits)) {
  split <- cv_folds_continuous$splits[[i]]

  # Extract training and validation data for this fold
  train_split <- training(split)
  valid_split <- testing(split)

  # Fit a GAM with sparsity
  gam_model <- gam(MIR ~ s(Hrs_Below_50_lag) + s(Hrs_Below_50_current) + 
                     s(Avg_HourlyWetBulbTemperature_current) + s(Avg_HourlyWindSpeed_current) + 
                     s(MEL_Prop) + s(average_daylight_hours) + s(Average_MEL),
                   data = train_split, method = "REML", select = TRUE)

  # Make predictions on the validation set
  preds <- predict(gam_model, newdata = valid_split, type = "response")

  # Create a tibble that includes both the actual values and predictions
  results_tibble <- tibble(
    truth = valid_split$MIR,
    estimate = preds
  )

  # Calculate RMSE for this fold using the results tibble
  cv_results[[i]] <- rmse(results_tibble, truth = truth, estimate = estimate)
}

# First, verify the structure of the first element to ensure we understand it correctly
print(cv_results[[1]])

# If cv_results[[1]] is indeed a tibble with an .estimate column, proceed with the following:
avg_rmse_GAMS <- mean(sapply(cv_results, function(x) x$.estimate))

# Display the average RMSE
cat("Average CV RMSE:", avg_rmse_GAMS, "\n")

# Fit the final GAM model on the full training data
final_gam_model <- gam(MIR ~ s(Hrs_Below_50_lag) + s(Hrs_Below_50_current) + 
                         s(Avg_HourlyWetBulbTemperature_current) + s(Avg_HourlyWindSpeed_current) + 
                         s(MEL_Prop) + s(average_daylight_hours) + s(Average_MEL),
                       data = train_data_continuous, method = "REML", select = TRUE)

# Create predictions for the training set using the final GAM model
train_preds <- predict(final_gam_model, newdata = train_data_continuous, type = "response")

# Create a tibble that includes both the actual values and predictions for the training set
train_results <- tibble(
  truth = train_data_continuous$MIR,  # Actual values
  estimate = train_preds  # Predicted values
)

# Calculate RMSE for the training set
train_rmse_result_GAMS <- rmse(train_results, truth = truth, estimate = estimate)

# Make predictions for the test set using the final GAM model
test_preds <- predict(final_gam_model, newdata = test_data_continuous, type = "response")

# Create a tibble that includes both the actual values and predictions for the test set
test_results <- tibble(
  truth = test_data_continuous$MIR,  # Actual values
  estimate = test_preds  # Predicted values
)

# Calculate RMSE for the test set
test_rmse_result_GAMS <- rmse(test_results, truth = truth, estimate = estimate)

# Display the RMSE for the training set
cat("Training Set RMSE:", train_rmse_result_GAMS$.estimate, "\n")
# Display the RMSE for the test set
cat("Test Set RMSE:", test_rmse_result_GAMS$.estimate, "\n")

# Calculate AIC for the final GAM model
aic_value <- AIC(final_gam_model)

# Display the AIC value
cat("AIC for the final GAM model:", aic_value, "\n")

```
# Gradient Boosting Machine (GBM)
```{r}

library(tidymodels)
library(dplyr)
library(yardstick)
library(ranger) # For Random Forest
library(boot) # For bootstrapping

# Load the dataset
mir_1part_all <- read_csv(here("data", "processed-data", "all_mir_clean_joined.csv"))

# Define predictors for the continuous outcome
predictors <- c("Hrs_Below_50_lag", "Hrs_Below_50_current", 
                "Avg_HourlyWetBulbTemperature_current", "Avg_HourlyWindSpeed_current", 
                "MEL_Prop", "average_daylight_hours", "Average_MEL")

# Prepare the dataset for continuous outcome
train_data_continuous <- mir_1part_all %>%
  select(MIR, all_of(predictors)) %>%
  na.omit()

# Split the data
set.seed(123)
data_split <- initial_split(train_data_continuous, prop = 0.75)
train_data_continuous <- training(data_split)
test_data_continuous <- testing(data_split)

# Define the tuning grid
tuning_grid <- grid_regular(
  trees(range = c(50, 500)), # Adjusted for GBM
  tree_depth(range = c(1, 5)), # Shallow trees
  learn_rate(range = c(0.01, 0.1)), # Learning rate for gradient descent
  min_n(range = c(5, 25)), # Adjusted based on preliminary insights
  levels = 10 # Increased resolution
)

# Define cross-validation folds
cv_folds_continuous <- vfold_cv(train_data_continuous, v = 5)

# Define the GBM model specification
gbm_model_spec_continuous <- boost_tree(trees = tune(), tree_depth = tune(), learn_rate = tune(), min_n = tune()) %>%
  set_engine("xgboost", objective = "reg:squarederror", nthread = parallel::detectCores()) %>%
  set_mode("regression")

# Create the recipe
gbm_recipe_continuous <- recipe(MIR ~ ., data = train_data_continuous)

# Workflow setup
gbm_workflow_continuous <- workflow() %>%
  add_recipe(gbm_recipe_continuous) %>%
  add_model(gbm_model_spec_continuous)

# Tune the model
tuning_results_continuous <- tune_grid(
  gbm_workflow_continuous,
  resamples = cv_folds_continuous,
  grid = tuning_grid,
  metrics = metric_set(yardstick::rmse)
)

# Extract the best hyperparameters
best_hps_continuous <- select_best(tuning_results_continuous, metric = "rmse")

# Fit the final model with the best hyperparameters on the full training dataset
final_fit_continuous <- fit(
  gbm_workflow_continuous %>% finalize_workflow(best_hps_continuous),
  data = train_data_continuous
)

### Calculate RMSE for the training and testing sets

# Predict on the training and testing set
train_preds <- predict(final_fit_continuous, new_data = train_data_continuous)$.pred
test_preds <- predict(final_fit_continuous, new_data = test_data_continuous)$.pred

# You need to create tibbles that combine the actual values with the predictions
train_results <- tibble(
  truth = train_data_continuous$MIR,
  estimate = train_preds
)

test_results <- tibble(
  truth = test_data_continuous$MIR,
  estimate = test_preds
)

# Now, you can correctly calculate RMSE for both the training and testing sets
train_rmse_result_gbm <- rmse(train_results, truth = truth, estimate = estimate)
test_rmse_result_gbm <- rmse(test_results, truth = truth, estimate = estimate)

# Display the RMSE results
cat("Training Set RMSE:", train_rmse_result_gbm$.estimate, "\n")
cat("Testing Set RMSE:", test_rmse_result_gbm$.estimate, "\n")

# Define a function to fit the model and calculate RMSE
boot_fit_and_evaluate <- function(data, indices) {
  # Extract the bootstrap sample
  bootstrap_data <- data[indices, ]

  # Fit the model on the bootstrap sample
  model <- gbm(MIR ~ ., data = bootstrap_data, distribution = "gaussian",
               n.trees = 1000, interaction.depth = 4, shrinkage = 0.01)

  # Predict on the original dataset
  preds <- predict(model, newdata = data, n.trees = 1000)

  # Compute RMSE
  rmse <- sqrt(mean((preds - data$MIR)^2))

  # Return RMSE
  return(rmse)
}

# Continue with bootstrapping to evaluate model's robustness
set.seed(123) # for reproducibility
bootstrap_results <- boot(
  data = train_data_continuous,
  statistic = boot_fit_and_evaluate,
  R = 100 # number of bootstrap resamples
)

# Calculate the bootstrapped RMSE
mean_rmse <- mean(bootstrap_results$t)
sd_rmse <- sd(bootstrap_results$t)
ci_rmse <- boot.ci(bootstrap_results, type = "perc")  # Percentile interval

# Extracting the correct CI values for percentile CI
lower_bound <- ci_rmse$percent[4]
upper_bound <- ci_rmse$percent[5]

# Output the bootstrapping results
cat("Bootstrapped RMSE: Mean=", mean_rmse, ", SD=", sd_rmse, "\n")
cat("95% CI for RMSE (Percentile): ", lower_bound, "-", upper_bound, "\n")

autoplot(tuning_results_continuous)

```

# SUmmarize the RSME values
```{r}
rmse_table <- data.frame(
  Model = c("Null Model", "Generalized Additive Model", "Random Forest", "LASSO",
            "Generalized Additive Model with Sparsity", "Gradient Boosting Machine"),
  Train_RMSE = c(null_rmse, rmse_train_gam1, train_rmse_result_rf$.estimate, 
                 train_rmse_result_lasso$.estimate, train_rmse_result_GAMS$.estimate, 
                 train_rmse_result_gbm$.estimate),
  Test_RMSE = c(null_rmse, rmse_test_gam1, test_rmse_result_rf$.estimate, 
                test_rmse_result_lasso$.estimate, test_rmse_result_GAMS$.estimate, 
                test_rmse_result_gbm$.estimate)
)
# Sort the table by Test_RMSE column in descending order
rmse_table <- rmse_table[order(-rmse_table$Test_RMSE), ]

# Print the sorted table
print(rmse_table)

#save as an RDS file
saveRDS(rmse_table, here("results", "tables", "rmse_table.rds"))

```