### 2024-04-25 products/manuscript/Manuscript.qmd

---
title: "Factors Influencing Abundance & Virus Isolation in Mosquitoes"
subtitle: "Which environmental, ecological, and temporal factors impact mosquito populations"
author: Andrew Ruiz
date: "`r Sys.Date()`"
md_number: true
format:
  docx:
    toc: false
    highlight-style: github
    reference_docx: path/to/your/reference/docx/file.docx
editor: 
  markdown: 
    wrap: 72
bibliography: 
  - ../../assets/project_mada.bib
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(dplyr)
library(readr)

here::set_here()

```

::: {#summary .unnumbered}
# Summary/Abstract

Eastern equine encephalitis (EEE) is a rare but serious disease caused by an Alphavirus transmitted to humans by infected mosquitoes. With a fatality rate of 30-50% and potential lifelong disability for survivors, understanding its transmission cycle is crucial. EEE transmission involves two cycles: an enzootic cycle between Culiseta melanura mosquitoes and birds, and an epizootic cycle where bridge vector mosquitoes transmit the virus to humans and other mammals. While C. melanura poses minimal direct threat to humans, isolations of EEE virus in these mosquitoes serve as early indicators of virus circulation. Conversely, isolations in bridge vector species signal heightened transmission risk. This research aims to identify significant relationships between variables to inform dissertation prospectus development and guide methodological approaches.
:::

# Problem statement

Which environmental,ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts and which
models best describe the relationship between these factors?

# Introduction

## Background

Eastern equine encephalitis (EEE) is caused by an Alphavirus transmitted
to humans by the bite of an infected mosquito. Human infections are
rare, but serious [@lindseyEasternEquineEncephalitis2018]. Humans and
other mammals are dead-end hosts for EEE. About 94% of human EEE
infections are asymptomatic. While symptomatic cases are rare, the
consequences can be severe. Mortality rates associated with symptomatic
infections range from 30 to 50% [@morensEasternEquineEncephalitis2019].
Those who survive often experience lifelong disability that includes
debilitating neurological damage
[@massachusettsdepartmentofpublichealthArbovirusSurveillancePlan2024].
On average, there are 11 cases reported in the US annually. Most occur
in Massachusetts and Florida [@vanderkelenRiskIndexModel2014]. Until
recently, nearly all human EEE cases in Massachusetts occurred in two
counties: Bristol and Plymouth.

Since 2000, there have been 45 human cases recorded in Massachusetts,
resulting in 22 deaths
[@massachusettsdepartmentofpublichealthArbovirusSurveillancePlan2024].
There are no human vaccines for EEE, and treatment consists of
palliative care only. Prevention measures include personal behaviors to
avoid mosquito bites and decreasing the mosquito population through
pesticide use and environmental modifications like removing standing
water
[@centersfordiseasecontrolandpreventionPreventionEasternEquine2022].

The Bristol County Mosquito Control Project (BCMCP) coordinates mosquito
surveillance and testing in the county from June to October. BCMCP has
used the same sentinel collection sites and trapping methods for over 40
years. Once trapped, mosquitoes are sorted and counted by species, and
vector species are submitted to the Massachusetts Department of Health’s
(MDPH) State Lab for PCR virus testing. Results are available within 24
hours of submission. When rates of EEE mosquito infections are above a
defined threshold, MDPH notifies the local boards of health and
recommends preventive measures. Recommended measures include outreach
and education to increase personal protective practices . Additionally,
MDPH may recommend that cities and towns in the affected areas cancel
evening outdoor events and discourage outdoor activities when mosquito
vectors are most active (dawn and dusk). When infection rates are high
enough to indicate imminent human transmission, the Commonwealth of
Massachusetts will recommend and fund aerial pesticide applications over
the affected areas.

## Study area

Bristol County, Massachusetts covers an area of 691 square miles with a
population density of 962 people per square mile. It borders Rhode
Island to the west, Plymouth County, MA to the east, and Norfolk County,
MA to the north. Naushon Island and Martha's Vineyard are across
Buzzard's Bay to the Southeast.

Atlantic White Cedar (AWC) swamps span over 4,300 acres within
Massachusetts, primarily concentrated in Plymouth and Bristol counties.
These swamps serve as critical breeding habitats for *Culiseta melanura*
(MEL)
[@commonwealthofmassachusettsdivisionoffisheries-wildlifeNaturalCommunityFact2007],
which is recognized as the primary enzootic vector for Eastern Equine
Encephalitis (EEE). AWC trees have large root systems that form "crypts"
and hold water throughout the year. These crypts serve as a main habitat
for MEL larval growth and provide shelter during winter months.

Between 2003 and 2022 Plymouth and Bristol counties were the epicenters
for EEE transmission in Massachusetts, accounting for 24 out of the
state's 41 human EEE cases.The Bristol County Mosquito Control Project
(BCMCP) conducts mosquito surveillance and abatement activities for the
county and often coordinates activities with Plymouth County Mosquito
Control Project.

## Questions/Hypotheses to be addressed
**Which environmental,ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts and which
models best describe the relationship between these factors?** 
From prior
research we know that the mosquito lifecyle is influenced by a number
of factors. Temperature plays a major role in the mosquito life cyle
virus transmission dynamics. As temperature increases, the rate of
larval development also increases and the time between when a mosquito
ingests virus in a bloodmeal and when it can transmit it decreases. The
result is that the mosquito population can grow at faster rate and they
will have more time to transmit the virus. However, MEL larvae develop
in well-insulated crypts that are less sensitive to ambient air
temperature changes, so the effect of temperature changes on MEL
development speed may be less pronounced than other vector species that
develop in open, standing water, like ditches and containers.
Precipitation increases the amount of standing water and increases the
chances of larval survival during the mosquito season and also during
the winter months for those species that overwinter in the larval stage.
For EEE, the main enzootic vector, MEL, overwinters as larvae.

While these relationships are well known, they are not well described
for EEE, especially in terms of quantifying the role these two factors
play in EEE transmission. Specifically, there's a gap in the
understanding of how these factors influence the number of EEE vectors mosquitoes
collected in traps and the rate of virus infection among those mosquitoes.

## Data collection and processing

### Mosquito surveillance data

The mosquito surveillance dataset encompasses counts by species from all
mosquito traps deployed in Bristol County, MA, from 2007 to 2023. The
raw dataset includes over 45,000 records and includes temporal and
spatial variable. The BCMCP, perform mosquito surveillance from June
through October annually.

During this interval, CDC light traps baited with carbon dioxide are
placed at various sentinel sites each week. Designed to attract
mosquitoes in search of blood meals, light traps serve as a reliable
measure of the actively feeding mosquito population, thereby
establishing them as the standard method for EEE surveillance. Some of
the sentinel site have been used for over 50 years.

Additionally, the county employs gravid traps, which are specifically
aimed at attracting and capturing gravid *Culex* mosquitoes looking for
egg-laying sites. While these traps are primarily used for West Nile
Virus (WNV) surveillance, they hold significant value for EEE monitoring
as well. Given the role of *Culex* species as bridge vectors for EEE,
their capture in gravid traps provides critical insight into the risk of
virus transmission.

The traps are installed in the afternoon and retrieved the following
morning. After the traps are collected, the next are stored in a freezer
to preserve and kill the mosquitoes. The mosquitoes are sorted by
species using a dissecting scope. Counts for each species are recorded
in the database and a portion of the vector species are driven to the
Massachusetts State Laboratory in Boston. The lab tests each mosquito
pool (or batch) for WNV and EEEv using PCR. Results are shared with
BCMCP on the same day of testing. Virus isolation results are recorded
in the collection records.

In addition to mosquito counts by species, the dataset includes the town
or city in which they were collected, the collection date, and whether
the mosquitoes were submitted for virus testing.

### Minimun infection rate (MIR)

One indicator of the level of virus activity present in the mosquito
population can be calculated from the information contained in the
mosquito dataset. The minimum infection rate (MIR) can be calculated at
various temporal and spatial scales from trap site over one day to all
sites over an entire season. MIR is calculated as: $$
MIR = \left( \frac{\text{Number of positive mosquito pools}}{\text{Total number of mosquitoes tested}} \right) \times 1000
$$ MIR provides an estimate of the prevalence of a specific pathogen
within a mosquito population. While it does not directly measure the
risk to humans, a higher MIR in mosquitoes that commonly bite humans can
indicate a higher risk of transmission. A study pubulihsed in 2007
indicated that in years with human cases in Massachusetts between 1979
and 2004, MIR was ≥1.0 by week 31 [@hachiyaHumanEasternEquine2007]. MIR
allows for the comparison of infection rates across different
geographical areas or time periods. This is crucial for identifying
areas or times of higher risk and for evaluating the effectiveness of
control measures.

![Minimum infection rate by
week](images/MIR_by_week_heatmap.png){fig-align="center"}

The surveillance dataset is structured to conform with state and federal
reporting requirements. As a result, the format of the data must be
transformed to in order to make it ready for analysis. Once the data
structure is corrected, it can be joined to other relevant datasets
using spatial or temporal attributes.

### Weather data

Weather plays a significant role in the lifecyle of the mosquito and
influences the rate at which a pathogen can develop in a mosquito.
Hourly and weather summaries were downloaded from NOAA's Climate Data
Portal. The datasets cover the years 2007-2023 at the Taunton Airport in
Bristol County.

The hourly summaries were used to calculate weather variables, such as
the daily maximum temperature and precipitation. While daily summaries
provide this information will less processing, the hourly summaries
provide more granular data that can be used to calculate additional
variables, such as the number of hours above a certain temperature
threshold or degree-days. Degree-days is a measure of the amount of heat
accumulated over a period of time. It is calculated by summing the
difference between the average daily temperature and a base temperature.
Degree-days are used to estimate the rate of development of insects and
other organisms that are sensitive to temperature. For mosquitoes,
degree-days can be used to estimate the rate of larval development and
the time between when a mosquito ingests a virus and when it can
transmit it [@guStatisticalEstimationDegree2006]. To simplify this
variable, the number of hours the temperature was ≥50°F and ≤86°F. These
thresholds were chosen based on the known temperature range for mosquito
development and virus transmission.

In addition to the weather summaries, drought data will be included in
the models. West Nile Virus (WNV) research suggests that drought has
been a driver in WNV outbreaks. While the lack of water can lead to
smaller mosquito populations, drought can lead to higher prevelance of
mosquito infection [@paullDroughtImmunityDetermine2017]. Drought data
was obtained from the US Drought Monitor and the Federal Emergency
Management Agency (FEMA). The data includes weekly drought severity
ratings for the entire United States. The data is available at county
and regional levels and will be used to calculate the number of weeks in
drought conditions and the severity of the drought. For this project,
drought data for Bristol County, MA, will be used and incorporated with
FEMA East Coast regions (I-IV) drought indicators.

## Data aquisition

With the exception of the mosquito collection data, all other datasets
are publicly available for use without restrictions.

## Data import and cleaning

Datasets will be loaded into R in a Quarto document to document and the
processing methods will be documented.

The Massachusetts Department of Public Health requires that all mosquito
control districts submit their collection data in a specific format.
This format divides each trap event into separate rows based on mosquito
species and whether they were submitted for virus testing. However, this
format is not useful for certain calculations where a zero count for a
species is not explicitly recorded. For example, even when sorted by
species to calculate the average count per trap, the average could be
inflated since the records will not include trap events with a zero
count of that species. In order to correct this, the data was pivoted to
create a column for every unique "species code" then the "pool size" for
each species code is summed for each trap event. A trap event is when
"town", "date of collection", "trap type" are all the same. For species
codes not included in the orginal table for a trap event, the field will
be assigned a zero value. The resulting table will have one row for
every trap event.

# Methods

Symptomatic EEE, and to a lesser extent WNV, human cases are rare. On
average, the United States records only about 9 cases of EEE annually.
Notably, Massachusetts reports more cases than any other state except
Florida, highlighting regional disparities in disease occurrence.
However, the incidence rate for neuroinvasive EEE cases in Bristol
County, Massachusetts, is estimated to be only around 0.06 per 100,000
people
[@centersfordiseasecontrolandpreventionPreventionEasternEquine2022]. The
low incidence of human cases poses significant challenges for using
traditional statistical models, making it difficult to accurately
estimate human transmission risk. This difficulty is largely due to the
sparse nature of the data and the variability in transmission patterns,
which complicates efforts to predict outbreaks and necessitates
alternative modeling approaches.

MIR will be the main outcome of interest for this project. MIR values
will be calculated for each week during the surveillance season over the
entire county. To reduce bias, only mosquitoes captured using carbon
dioxide-baited CDC light traps will be used to to calculate MIR. These
traps are the standard EEE surveillance and are used to monitor the
actively feeding mosquito population. Additionally, these trap types
have been used in this area for over 40 years, many at the same sites
currently in use.

After MIR was calculated and joined with weather and climate data, the
distribution of the MIR was examined more closely. About 72% of the
records were zero values, which is not uncommon for this type of data.
The non-zero values were skewed to the left and there was strong
evidence of over-dispersion.

![](images/mir_density_plot.png){fig-align="center"}

The average MIR was 1.26, with an observed variance of 7.56, resulting
in a Variance-to-Mean Ratio (VMR) of 6.00. This pronounced
overdispersion suggests that mosquito infection rates were not uniformly
distributed over time or space. Instead, they may have been influenced
by a range of factors, such as environmental conditions or variations in
amplifying host abundance.

The data was visualized to identify trends and relationships between the
variables. There were no clear patterns in the data between MIR and the
potential predictors. However, several varibles showed a non-linear
relationship with MIR. For example, the relationship between MIR and
average temperature was not linear. Instead, MIR increased with
temperature up to a certain point, then decreased. This suggests that
the relationship between temperature and MIR may be more complex than a
simple linear model can capture.

In response to the observed non-linear relationships in the data, a
two-stage Generalized Additive Model (GAM) analysis was used. GAMs
provide a versatile framework for capturing complex, non-linear
relationships between response variables and multiple predictors, which
are often encountered in ecological data sets.

In the first stage of the analysis, the GAM was utilized to address the
high incidence of zeros in the dataset—an indicator of overdispersion.
This zero-inflated model allowed us to separately model the presence
versus absence of the outcome variable, thereby appropriately accounting
for the excessive zeros.

Subsequently, the second stage of the GAM focused on modeling the
continuous part of the MIR data, relating it to the suite of potential
predictors. This step incorporated smooth functions of the predictors,
permitting the data to inform the shape of the relationship without
imposing a predetermined form, as would be the case in a traditional
linear model.

While the two-stage GAM appeared to be a good fit, there were too few
observations to provide a robust estimate of the relationship between
the predictors and MIR. The original dataset contained over 40,000
observations, but after reformatting the data to fit a rectangular shape
needed for analysis, there were less than 400 observations in the final
set. After the zeros were removed for part 2 of the GAM, the sample
number of observations left was less than 100. Because of this, it was
decided to use a single GAM with a dataset that included zero and
non-zero observations. Additionally, all trap types were included in the
analysis, not just the carbon dioxide-baited CDC light traps. While some
precision was lost when the one part model was adopted, the gain in
observations was deemed more valuable.

### Schematic of workflow

*Schematic will be refined in later steps. The doc will also be
reformatted to move the schematic to an appendix* *The image should
appear upon rendering*

```{mermaid}
flowchart TB
    id1["Questions"] --> id2["Data Collection"]
    id2 --> id3["Data Loading and Cleaning"]
    id3 --> id4["Data Cleaning"]
    id4 --> id5["Data Transformation"]
    id5 --> id6["Data Exploration & Visualization"]
    id6 --> id7["Data Modeling"]
    id7 --> id8["Evaluation"]

    id8 -.-> id1
    id8 --> id9[Interpretation]
    id9 --> id10[Communication]

    classDef mainProcess fill:#fff,stroke:#333,stroke-width:2px;

    class id1,id2,id3,id4,id5,id6,id7,id8,id9,id10 mainProcess
```

## Exploratory/Descriptive analysis
One factor to consider is if early virus isolations are associated with
cases in humans and other mammals. The table below shows the date of the
first EEE virus isolation in mosquitoes and the number of human of other
animal cases by year.

![](images/eee_isolation_cases-04.png){fig-align="center" width="251"}

*Culiseta melanura* (MEL) plays an important role in EEE virus
amplification in the enzootic cycle. Understanding the population
dynamics might provide more insight into what is driving the spikes,
like in 2019.

The graph below shows the proportion of MEL among all mosquito vector
species trapped in Bristol County, MA.

![](images/mel_proportion.png){fig-align="center"}

From this we see a large proportion of MEL in the traps for 2019. In
fact, they comprise more than half of all vector species for that year.
Another interesting aspect is the high proportion in 2018. MEL
overwinters as larvae and as long as there is sufficient standing water
to last the winter, MEL hatched in the previous fall would survive to
emerge the next year. This suggests it might be important to look at
conditions in the previous year or season for a better understanding.

Since water plays a significant role in the mosquito life cycle, let's
look at the annual precipitation.

![](images/annual_precip-02.png){fig-align="center"}

The graph indicates that the annual precipitation total for 2018 is
higher than the other years. This suggest that there was enough water to
support large MEL larval populations though out the winter.

Another view plots the average annual MEL catch against total
precipitation.

![](images/combined_trends_mel_precip.png){fig-align="center"}

## Basic statistical analysis
Basic linear models assessed environmental factors' impact on the EEE virus transmission cycle. Initial models investigated the link between virus isolation in mosquitoes and weather conditions (average temperature by week and total precipitation). The outcome was whether at least one positive test occurred for a pool in a given week. The predictor in the first model was the average maximum daily temperature from the previous week. The logistic model revealed a significant relationship. Higher temperatures were positively associated with the likelihood of a positive result (odds ratio ≈ 1.069 for each unit increase in temperature; Estimate = 0.06695, SE = 0.02033, p < 0.001). Incorporating Lagged_Avg_TMAX as a predictor improved the model, reducing residual deviance to 405.79 from the null deviance of 417.24, indicating a better fit. However, residual deviance on 299 degrees of freedom suggests unexplained variability, indicating other influential factors. The model's AIC of 411.59 suggests a balance between complexity and predictive power, supporting its utility. Yet, this relationship warrants closer examination due to likely autocorrelation of temperature with progressing weeks during summer months

While
precipitation has been demonstrated to influence the population dynamics
it did not show to have any effect on a positive result. This was
initailly surprising, but upon after considering the specific habitat of
the main EEE enzootic vector, *Culiseta melanura*, it makes sense. MEL
larvae develop in well insulated crypts that are less sensitive
evaporative processes than surface water. The effect of precipitation on
MEL larval habitat abundance may be less pronounced than other vector
species that develop in open, standing water, like ditches and
containers. Another model assessed whether the temperature precipitation
in April, May, and June of each year effected the date of the first
positive pool test (R file: PT3_edacode). The date of the first positive
pool was identified and assigned a day number from January 1st of that
year. The predictors in this linear model were the average maximum
temperature for April, May, and June and the sum of precipitation for
the same time period. None of these predictors were significant.

There was one other significant result. When MIR is set as the
dependent, continuous variable in a linear regression and the average
maximum temperature for the previous week is used as the predictor, the
model suggests there is a significant inverse relationship. As average
temperature increases, the MIR decreases (model_mir_temp.R). This
relationship is counterintutive, especially given the result of the
logistic regression and requires more examination.

## Statistical analysis
Before any models were assessed, the RSME for the null model was calculated (2.40). This was used as the basis for assessing model improvement.

Analysis began with a Generalized Additive Model (GAM) with predictors that were selected based on the exploratory single variable GAM. The variables were chosen if their p=values were <0.05. Eight variables met that condition including: the number of hours below 50°F the week before the mosquito collection week (Hrs_Below_50_lag), the number of hours below 50°F the week of collection (Hrs_Below_50_current), the drought severity index for East Coast FEMA regions (DSCI_fema), the average hourly relative humidity the week of collection (Avg_HourlyRelativeHumidity_current), the average hourly windspeed the week of collection (Avg_HourlyWindSpeed_current), the average number of daylight hours per day (average_daylight_hours), the average number of MEL captured by trap (Average_MEL), and the proportion of MEL captured out of the total number of mosquitoes trapped (MEL_Prop).

All models included training (.75) and a testing (.25) dataset that were generated using the random seed 123. All analysis can be found in the analysis.qmd file.  The first model was a GAM with all eight predictors, followed by a Random Forest, LASSO, Generalized Additive Model with Sparsity (GAMS), and a Gradient Boosting Machine (GBM) model.

As expected, the RSME is lower for the training dataset than the testing dataset for all models. The RMSE values are displayed in the table below.

```{r}
# Define the file path using here()
file_path_rds <- here("results", "tables", "rmse_table.rds")

# Read the RDS file
rmse_table <- readRDS(file_path_rds)
rmse_table
```
# Discussion
The null model RMSE was 2.40. All models showed improvement over the null model on both the training and testing datasets. The GAM shows significant improvement over the null model for the training set, but only slightly improves over the null for the testing set. This suggests that beyond capturing the mean response, this GAM is not well generalized. The substantial improvement in the training set RSME for GBM and Random Forest coupled with less improvement over the null for the testing set, suggest a tendency for overfitting. These models may be capturing the training data's patterns that do not gerneralize well to unseen data. The the gap between the training and testing RMSE for the LASSO and GAMS models suggest there is less overfitting and are slightly better at generalizing than the others. However, given the non-linear relationships, the LASSO model is likely not appropriate for this analysis.

## Strengths and Limitations

The accuracy
and reliability of the mosquito collection data is excellent. The same
procedures and sentinel sites have were in place for the entire
analytical horizon. Further, there were only two people responsible for
sorting, identifying and counting species over the entire period. However, the data is not without limitations. Mosquito surveillance is inherently biased. Traps are set in locations that are most likely to capture infected mosquitoes. While this is a positive for early warning of virus activity, it is not necessarily representative of the surrounding area. Probably the biggest limitation is the small number of observations. The mosquito season is short in Massachusetts and funding for trapping and testing mosquitoes is limited. Mosquito control efforts, like pesticide applications, are missing from this analysis. These efforts are designed to lower the levels of virus transmitting mosquitoes and would likely provide more insight into the relationship between the predictors and virus activity.

MIR is not a perfect measure of virus activity in mosquitoes. It tends
to underestimate the level of virus circulating in the environment.
However, other measures require equal pool numbers for testing. There is
a correction factor for variable pool numbers and I will explore this
option in later phases of this project. Ideally, the outcome would be human or animal cases. However, these cases are rare and modeling them given the limited amount of surveillance data may not be possible.

Initially, the lack of significance for most temperature and precipitation variable was surprising. However, upon closer examination, it makes sense. The main enzootic vector of EEE and the one targeted for surveillance, *Culiseta melanura*, develops in well insulated crypts that are less sensitive to ambient air temperature changes. The effect of temperature changes on MEL development speed may be less pronounced than other vector species that develop in open, standing water, like ditches and containers. Precipitation increases the amount of standing water and increases the chances of larval survival during the mosquito season and also during the winter months for those species that overwinter in the larval stage. For EEE, the main enzootic vector, MEL, overwinters as larvae.

The spatial scale is an issue with the human and veterinary cases data.
Human case data is released only on a county scale to ensure patient
privacy. Some veterinary case data is available on the town-level, but
most remains only on the county-level. These limitations are barriers to
accurate spatial analysis. The project will rely on virus isolation in
mosquitoes to model risk. The available bird data relies heavily on
crowd-sourced reports. There is one standardized survey (Breeding Bird
Survey) that seems to be more accurate and reliable. However, this data
is not available for every year.

Likely the biggest limitation 
## Conclusions 
While all models demonstrated improvement over null model, none of them were robust enough to be used for reliable prediction. In the next phases of this project, the data will be expanded to include more years and cover a larger portion of Massachusetts. In additon to increasing the the number of observations, other sources of data will be incorporated. Including virus activity from surrounding regions and areas from which birds migrate to Massachusetts.

# References