### 2024-04-29 data_dictionary/data_dict_search_tool.qmd
Debug and clean
debugged code and updated README files

---
title: "Data Dictionary Search Tool"
output: html_document
runtime: shiny
editor: visual
---

# Data dictionary search tool

## Use this tool to search the data dictionary by file name, column name, or both.

### First code block was used to compile the data dictionary from multiple CSV files. The data dictionary is then saved as a CSV file named "Data_Dictionary.csv".

```{r echo=FALSE}
library(shiny)
library(here)
library(dplyr)
library(readr)
```

## Run this code this code to open the search tool.

### you can also view the entire data dictionary here: (.../data_dictionary/Data_Dictionary.csv)

```{r echo=FALSE}
# Load your data dictionary
data_dictionary <- read_csv(here("data_dictionary", "Data_Dictionary.csv"))

# Define UI
ui <- fluidPage(
    titlePanel("Data Dictionary Search Tool"),
    sidebarLayout(
        sidebarPanel(
            helpText("Use the fields below to search the data dictionary by file name, column name, or both."),
            textInput("file_name", "Enter File Name:", ""),
            textInput("column_name", "Enter Column Name:", ""),
            helpText("Press search to update the results."),
            actionButton("search", "Search"),
            helpText("Tip: The search is case-insensitive and partial matches will also show up in the results.")
        ),
        mainPanel(
            dataTableOutput("results")
        )
    )
)

# Define server logic
server <- function(input, output) {
    # Reactive value for filtered data
    filtered_data <- reactive({
        req(input$search)  # Make sure the search button has been pressed
        data_dictionary %>%
            filter(
                (input$file_name == "" | grepl(input$file_name, FileName, ignore.case = TRUE)) &
                (input$column_name == "" | grepl(input$column_name, ColumnName, ignore.case = TRUE))
            )
    })

    # Output the results
    output$results <- renderDataTable({
        filtered_data()
    })
}

# Run the application
shinyApp(ui = ui, server = server)

```