### 2024-04-30 products/manuscript/supplement/Supplementary-Material.qmd

---
title: "Supplement to Factors Influencing Abundance & Virus Isolation in Mosquitoes"

subtitle: "Schematic of workflow"
author: Andrew Ruiz
date: "2024-04-30"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
---
#### *The image should appear upon rendering*
```{mermaid width=700 height=500}
flowchart TB
    id1["Research Questions"] --> id2["Data Collection"]
    id2 --> id3["Data Loading"]
    id3 --> id4["Data Cleaning & Preprocessing"]
    id4 --> id5["Data Exploration & Visualization"]
    id5 --> id6["Variable Selection"]
    id6 --> id7["Model Development"]
    id7 --> id7a["Model Training (Random Seed 123)"]
    id7a --> id7b["Model Comparison\n(GAM, RF, LASSO, GAMS, GBM)"]
    id7b --> id8["Model Evaluation\n(RMSE Calculation)"]
    id8 -.-> id1
    id8 --> id9["Result Interpretation"]

    classDef mainProcess fill:#fff,stroke:#333,stroke-width:2px;
    class id1,id2,id3,id4,id5,id6,id7,id7a,id7b,id8,id9 mainProcess
```