### 2024-05-03 products/manuscript/Manuscript.qmd

---
title: "Factors Influencing Abundance & Virus Isolation in Mosquitoes"
subtitle: "Which environmental, ecological, and temporal factors impact mosquito populations"
author: Andrew Ruiz
date: "`r Sys.Date()`"
md_number: true
format:
  docx:
    toc: false
    highlight-style: github
    reference_docx: path/to/your/reference/docx/file.docx
editor: 
  markdown: 
    wrap: 72
bibliography: 
  - ../../assets/project_mada.bib
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(dplyr)
library(readr)

here::set_here()

```

::: {#summary .unnumbered}
# Summary/Abstract

Eastern equine encephalitis (EEE) is a rare but severe disease caused
by an Alphavirus transmitted to humans by infected mosquitoes. With a
fatality rate of 30-50% and potential lifelong disability for survivors,
understanding its transmission cycle is crucial. EEE transmission
involves two cycles: an enzootic cycle between Culiseta melanura
mosquitoes and birds and an epizootic cycle where bridge vector
mosquitoes transmit the virus to humans and other mammals. While *C. melanura* poses minimal direct threat to humans, isolations of EEE virus in these mosquitoes serve as early indicators of virus circulation. In contrast, isolations in bridge vector species signal a heightened transmission risk. This project aims to identify significant relationships between variables to inform the development of a dissertation prospectus and guide methodological approaches..
:::

# Research Question

In southeastern Massachusetts, the transmission dynamics of Eastern Equine Encephalitis (EEE) remain inadequately understood, particularly how environmental, ecological, and temporal factors interplay to affect outbreak patterns. This gap in knowledge hinders effective predictive modeling and response strategies, crucial for preventing and mitigating the impact of EEE.

**Which environmental,ecological, and temporal factors influence the transmission cycles for EEE in southeastern Massachusetts, and which models best describe the relationship between these factors?**

# Introduction

## Background

Eastern equine encephalitis (EEE) is caused by an Alphavirus transmitted
to humans by the bite of an infected mosquito. Human infections are
rare but serious [@lindseyEasternEquineEncephalitis2018]. Humans and
other mammals are dead-end hosts for EEE. About 94% of human EEE
infections are asymptomatic. While symptomatic cases are rare, the
consequences can be severe. Mortality rates associated with symptomatic
cases range from 30 to 50% [@morensEasternEquineEncephalitis2019].
Those who survive often experience lifelong disability that includes
debilitating neurological damage
[@massachusettsdepartmentofpublichealthArbovirusSurveillancePlan2024].
On average, 11 cases reported in the US annually. Most occur
in Massachusetts and Florida [@vanderkelenRiskIndexModel2014]. Until
recently, nearly all human EEE cases in Massachusetts occurred in two
counties: Bristol and Plymouth.

Since 2000, there have been 45 human cases recorded in Massachusetts,
resulting in 22 deaths
[@massachusettsdepartmentofpublichealthArbovirusSurveillancePlan2024].
There are no human vaccines for EEE, and treatment consists of
palliative care only. Prevention measures include personal behaviors to
avoid mosquito bites and decreasing the mosquito population through
pesticide use and environmental modifications like removing standing
water
[@centersfordiseasecontrolandpreventionPreventionEasternEquine2022].

The Bristol County Mosquito Control Project (BCMCP) coordinates mosquito
surveillance and testing in the county from June to October. BCMCP has
used the same sentinel collection sites and trapping methods for over 40
years. Once trapped, mosquitoes are sorted and counted by species, and
vector species are submitted to the Massachusetts Department of Health’s
(MDPH) State Lab for PCR virus testing. Results are available within 24
hours of submission. When rates of EEE mosquito infections are above a
defined threshold, MDPH notifies the local boards of health and
recommends preventive measures. Recommended measures include outreach
and education to increase personal protective practices. Additionally,
MDPH may recommend that cities and towns in the affected areas cancel
evening outdoor events and discourage outdoor activities when mosquito
vectors are most active (dawn and dusk). When infection rates are high
enough to indicate imminent human transmission, the Commonwealth of
Massachusetts will recommend and fund aerial pesticide applications over
the affected areas.

## Study area

Bristol County, Massachusetts covers 691 square miles with a population
density of 962 people per square mile. It borders Rhode Island to the
west, Plymouth County, MA to the east, and Norfolk County, MA to the
north. Naushon Island and Martha's Vineyard are across Buzzard's Bay.

Atlantic White Cedar (AWC) swamps cover more than 4,300 acres across
Massachusetts, with the largest concentrations found in Plymouth and
Bristol counties. These swamps are crucial breeding grounds for Culiseta
melanura (MEL), which is identified as the primary enzootic vector for
Eastern Equine Encephalitis (EEE)
[@commonwealthofmassachusettsdivisionoffisheries-wildlifeNaturalCommunityFact2007].
Atlantic White Cedar (AWC) trees possess extensive root systems that
create "crypts" capable of retaining water year-round. These crypts are
crucial habitats for the larval growth of Culiseta melanura (MEL) and
also provide shelter during the winter.

Between 2003 and 2022 Plymouth and Bristol counties were the epicenters
for EEE transmission in Massachusetts, accounting for 24 out of the
state's 41 human EEE cases.The Bristol County Mosquito Control Project
(BCMCP) conducts mosquito surveillance and abatement activities for the
county and often coordinates activities with Plymouth County Mosquito
Control Project.

## Questions/Hypotheses to be addressed

*Which environmental, ecological, and temporal factors influence the transmission cycles for EEE in southeastern Massachusetts, and which models best describe the relationship between these factors?* From prior research, we know that the mosquito lifecycle is influenced by several factors. Temperature plays a significant role in the mosquito life cycle and virus transmission dynamics. As temperature increases, the rate of larval development also increases, and the time between when a mosquito ingests the virus in a blood meal and when it can transmit it decreases. The result is that the mosquito population can grow at a faster rate, and they will have more time to transmit the virus. However, MEL larvae develop in well-insulated crypts that are less sensitive to ambient air temperature changes, so the effect of temperature changes on MEL development speed may be less pronounced than other vector species that develop in open, standing water, like ditches and containers. Precipitation increases the amount of standing water and the chances of larval survival during the mosquito season and winter for those species that overwinter in the larval stage. For EEE, the primary enzootic vector, MEL, overwinters as larvae.
While these relationships are well known, they are not well described for EEE, especially in quantifying the role these two factors play in EEE transmission. Specifically, there’s a gap in understanding how these factors influence the number of EEE vector mosquitoes collected in traps and the rate of virus infection among those mosquitoes

## Data collection and processing

With the exception of the mosquito collection data, all other datasets
are publicly available for use without restrictions.

### Mosquito surveillance data

The mosquito surveillance dataset encompasses counts by species from all mosquito traps deployed in Bristol County, MA, from 2007 to 2023. The raw dataset includes over 45,000 records and includes temporal and spatial variables. The BCMCP performs mosquito surveillance from June through October annually.
During this interval, CDC light traps baited with carbon dioxide are placed at various sentinel sites weekly. Designed to attract mosquitoes in search of blood meals, light traps serve as a reliable measure of the actively feeding mosquito population, thereby establishing them as the standard method for EEE surveillance. Some of the sentinel sites have been used for over 50 years.
Additionally, the county employs gravid traps specifically aimed at attracting and capturing gravid Culex mosquitoes looking for egg-laying sites. While these traps are primarily used for West Nile Virus (WNV) surveillance, they also hold significant value for EEE monitoring. Given the role of Culex species as bridge vectors for EEE, their capture in gravid traps provides critical insight into the risk of virus transmission.
The traps are installed in the afternoon and retrieved the following morning. After the traps are collected, the next are stored in a freezer to preserve and kill the mosquitoes. The mosquitoes are sorted by species using a dissecting scope. Counts for each species are recorded in the database, and a portion of the vector species are driven to the Massachusetts State Laboratory in Boston. The lab tests each mosquito pool (or batch) for WNV and EEEv using PCR. Results are shared with BCMCP on the same day of testing. Virus isolation results are recorded in the collection records.
In addition to mosquito counts by species, the dataset includes the town or city in which they were collected, the collection date, and whether the mosquitoes were submitted for virus testing.

### Minimum infection rate (MIR)

In our investigation into the environmental, ecological, and temporal
factors that impact the transmission cycles of EEE in southeastern
Massachusetts, the Mosquito Infection Rate (MIR) emerges as a pivotal
indicator. MIR provides a quantifiable measure of virus prevalence
within mosquito populations, making it an essential metric for modeling
the dynamics of disease spread. By analyzing changes in MIR in response
to varying environmental conditions and mosquito behaviors, we can
enhance our understanding of the key drivers behind EEE transmission.
This understanding, in turn, guides the development of predictive models
that can accurately forecast periods of heightened transmission risk
based on observed or anticipated changes in the identified factors. The
minimum infection rate (MIR) can be calculated at various temporal and
spatial scales from trap site over one day to all sites over an entire
season. MIR is calculated as: $$
MIR = \left( \frac{\text{Number of positive mosquito pools}}{\text{Total number of mosquitoes tested}} \right) \times 1000
$$ While MIR does not directly measure the risk to humans, a higher MIR in mosquitoes that commonly bite humans can indicate a higher risk of transmission. A study published in 2007 indicated that in years with human cases in Massachusetts between 1979 and 2004, the MIR was ≥1.0 by week 31 (mid-late July) [@hachiyaHumanEasternEquine2007]. MIR allows for comparing infection rates across different geographical areas or time periods. This is crucial for identifying areas or times of higher risk and evaluating control measures' effectiveness.

![Minimum infection rate by
week](images/MIR_by_week_heatmap.png){fig-align="center"}

The surveillance dataset is structured to conform with state and federal reporting requirements. As a result, the data format must be transformed to make it ready for analysis. Once the data structure is corrected, it can be joined to other relevant datasets using spatial or temporal attributes.

### Weather data

Weather plays a significant role in the lifecyle of the mosquito and
influences the rate at which a pathogen can develop in a mosquito.
Hourly and weather summaries were downloaded from NOAA's Climate Data
Portal. The datasets cover the years 2007-2023 at the Taunton Airport in
Bristol County.

The hourly summaries were used to calculate weather variables, such as
the daily maximum temperature and precipitation. While daily summaries
provide this information will less processing, the hourly summaries
provide more granular data that can be used to calculate additional
variables, such as the number of hours above a certain temperature
threshold or degree-days. Degree-days is a measure of the amount of
heat accumulated over a period of time. It is calculated by summing the
difference between the average daily temperature and a base temperature.
Degree-days are used to estimate the rate of development of insects and
other organisms that are sensitive to temperature. For mosquitoes,
degree-days can be used to estimate the rate of larval development and
the time between when a mosquito ingests a virus and when it can
transmit it [@guStatisticalEstimationDegree2006]. To simplify this
variable, the number of hours the temperature was ≥50°F and ≤86°F. These
thresholds were chosen based on the known temperature range for mosquito
development and virus transmission.

In addition to the weather summaries, drought data will be included in
the models. West Nile Virus (WNV) research suggests that drought has
been a driver in WNV outbreaks. While the lack of water can lead to
smaller mosquito populations, drought can lead to higher prevelance of
mosquito infection [@paullDroughtImmunityDetermine2017]. Drought data
was obtained from the US Drought Monitor and the Federal Emergency
Management Agency (FEMA). The data includes weekly drought severity
ratings for the entire United States. The data is available at county
and regional levels and will be used to calculate the number of weeks in
drought conditions and the severity of the drought. For this project,
drought data for Bristol County, MA, will be used and incorporated with
FEMA East Coast regions (I-IV) drought indicators.

## Data import and cleaning

Datasets will be loaded into R in a Quarto document to document and the
processing methods will be documented.

The Massachusetts Department of Public Health mandates a specific format
for mosquito control districts to submit their collection data. In this
format, each trap event is divided into separate rows based on mosquito
species and whether they were submitted for virus testing. However, this
format poses limitations for certain calculations where a zero count for
a species is not explicitly recorded. For instance, when sorting data by
species to calculate the average count per trap, the absence of records
with a zero count could inflate the calculated average.

To address this issue, the data was transformed by pivoting it to create
a column for every unique "species code." Then, the "pool size" for each
species code was summed for every trap event, defined as instances where
"town," "date of collection," and "trap type" are identical. For species
codes not originally included in the table for a trap event, the
corresponding field was assigned a zero value. As a result, the
resulting table now contains one row for each trap event.

# Methods

Symptomatic EEE human cases are rare. On average, the United States
records only about 9 cases of EEE annually. Notably, Massachusetts
reports more cases than any other state except Florida, highlighting
regional disparities in disease occurrence. However, the incidence rate
for neuroinvasive EEE cases in Bristol County, Massachusetts, is
estimated to be only around 0.06 per 100,000 people
[@centersfordiseasecontrolandpreventionPreventionEasternEquine2022]. The
low incidence of human cases poses significant challenges for using
traditional statistical models, making it difficult to accurately
estimate human transmission risk. This difficulty is largely due to the
sparse nature of the data and the variability in transmission patterns,
which complicates efforts to predict outbreaks and necessitates
alternative modeling approaches.

MIR will be the main outcome of interest for this project. MIR values
will be calculated for each week during the surveillance season over the
entire county.

After MIR was calculated and joined with weather and climate data, the
distribution of the MIR was examined more closely. About 72% of the
records were zero values, which is not uncommon for this type of data.
The non-zero values were skewed to the left and there was strong
evidence of over-dispersion.

![](images/mir_density_plot.png){fig-align="center"}

The average MIR was 1.26, with an observed variance of 7.56, resulting
in a Variance-to-Mean Ratio (VMR) of 6.00. This pronounced
overdispersion suggests that mosquito infection rates were not uniformly
distributed over time or space. Instead, they may have been influenced
by a range of factors, such as environmental conditions or variations in
amplifying host abundance.

The data was visualized to identify trends and relationships between the
variables. There were no clear patterns in the data between MIR and the
potential predictors. However, several varibles showed a non-linear
relationship with MIR. For example, the relationship between MIR and
average temperature was not linear. Instead, MIR increased with
temperature up to a certain point, then decreased. This suggests that
the relationship between temperature and MIR may be more complex than a
simple linear model can capture.

In response to the observed non-linear relationships in the data, a
two-stage Generalized Additive Model (GAM) analysis was used. GAMs
provide a versatile framework for capturing complex, non-linear
relationships between response variables and multiple predictors, often encountered in ecological data sets.

In the first stage of the analysis, the GAM was utilized to address the
high incidence of zeros in the dataset. This zero-inflated model allowed
us to separately model the presence versus absence of the outcome
variable, thereby appropriately accounting for the excessive zeros.

The second stage of the GAM focused on modeling the continuous part of
the MIR data, relating it to the suite of potential predictors. This
step incorporated smooth functions of the predictors, permitting the
data to inform the shape of the relationship without imposing a
predetermined form, as would be the case in a traditional linear model.

While the two-stage Generalized Additive Model (GAM) initially seemed to
be a suitable approach, there were insufficient observations to yield a
robust estimate of the relationship between the predictors and Minimum
Infection Rate (MIR). Despite starting with a dataset containing over
40,000 observations, the transformation of the data into a rectangular
shape required for analysis resulted in fewer than 400 observations in
the final dataset. After removing the zeros for part 2 of the GAM, the
number of observations decreased to less than 100. Consequently, a
decision was made to utilize a single GAM with a dataset containing both
zero and non-zero observations. Although adopting a single-part model
led to some loss of precision, the increase in the number of
observations was deemed more advantageous.

## Exploratory/Descriptive analysis

One factor to consider is if early virus isolations are associated with
cases in humans and other mammals. The table below shows the date of the
first EEE virus isolation in mosquitoes and the number of human of other
animal cases by year.

![](images/eee_isolation_cases-04.png){fig-align="center" width="251"}

*Culiseta melanura* (MEL) plays an important role in EEE virus
amplification in the enzootic cycle. Understanding the population
dynamics might provide more insight into what is driving the spikes,
like in 2019.

The graph below shows the proportion of MEL among all mosquito vector
species trapped in Bristol County, MA.

![](images/mel_proportion.png){fig-align="center"}

From this we see a large proportion of MEL in the traps for 2019. In
fact, they comprise more than half of all vector species for that year.
Another interesting aspect is the high proportion in 2018. MEL
overwinters as larvae and as long as there is sufficient standing water
to last the winter, MEL hatched in the previous fall would survive to
emerge the next year. This suggests it might be important to look at
conditions in the previous year or season for a better understanding.

Since water plays a significant role in the mosquito life cycle, let's
look at the annual precipitation.

![](images/annual_precip-02.png){fig-align="center"}

The graph indicates that the annual precipitation total for 2018 is
higher than the other years. This suggest that there was enough water to
support large MEL larval populations though out the winter.

Another view plots the average annual MEL catch against total
precipitation.

![](images/combined_trends_mel_precip.png){fig-align="center"}

## Basic statistical analysis

Basic linear models assessed environmental factors' impact on the EEE
virus transmission cycle. Initial models investigated the link between
virus isolation in mosquitoes and weather conditions (average
temperature by week and total precipitation). The outcome was whether at
least one positive test occurred for a pool in a given week. The
predictor in the first model was the average maximum daily temperature
from the previous week. The logistic model revealed a significant
relationship. Higher temperatures were positively associated with the
likelihood of a positive result (odds ratio ≈ 1.069 for each unit
increase in temperature; Estimate = 0.06695, SE = 0.02033, p \< 0.001).
Incorporating Lagged_Avg_TMAX as a predictor improved the model,
reducing residual deviance to 405.79 from the null deviance of 417.24,
indicating a better fit. However, residual deviance on 299 degrees of
freedom suggests unexplained variability, indicating other influential
factors. The model's AIC of 411.59 suggests a balance between complexity
and predictive power, supporting its utility. Yet, this relationship
warrants closer examination due to likely autocorrelation of temperature
with progressing weeks during summer months

While precipitation has been demonstrated to influence the population
dynamics it did not show to have any effect on a positive result. This
was initailly surprising, but upon after considering the specific
habitat of the main EEE enzootic vector, *Culiseta melanura*, it makes
sense. MEL larvae develop in well insulated crypts that are less
sensitive evaporative processes than surface water. The effect of
precipitation on MEL larval habitat abundance may be less pronounced
than other vector species that develop in open, standing water, like
ditches and containers. Another model assessed whether the temperature
precipitation in April, May, and June of each year effected the date of
the first positive pool test (R file: PT3_edacode). The date of the
first positive pool was identified and assigned a day number from
January 1st of that year. The predictors in this linear model were the
average maximum temperature for April, May, and June and the sum of
precipitation for the same time period. None of these predictors were
significant.

There was one other significant result. When MIR is set as the
dependent, continuous variable in a linear regression and the average
maximum temperature for the previous week is used as the predictor, the
model suggests there is a significant inverse relationship. As average
temperature increases the week prior to collection, the MIR decreases
(model_mir_temp.R). This relationship is counterintutive, especially
given the result of the logistic regression and requires more
examination.

## Statistical analysis

All detailed model analyses are available in the analysis.qmd file,
providing a transparent account of the methodologies employed and the
statistical rigor behind the conclusions drawn.

Before any models were assessed, the RSME for the null model was
calculated (2.40). This was used as the basis for assessing model
improvement.

Analysis began with a Generalized Additive Model (GAM) with predictors
that were selected based on the exploratory single variable GAM. The
variables were chosen if their p=values were \<0.05. Eight variables met
that condition including: the number of hours below 50°F the week before
the mosquito collection week (Hrs_Below_50_lag), the number of hours
below 50°F the week of collection (Hrs_Below_50_current), the drought
severity index for East Coast FEMA regions (DSCI_fema), the average
hourly relative humidity the week of collection
(Avg_HourlyRelativeHumidity_current), the average hourly wind speed the
week of collection (Avg_HourlyWindSpeed_current), the average number of
daylight hours per day (average_daylight_hours), the average number of
MEL captured by trap (Average_MEL), and the proportion of MEL captured
out of the total number of mosquitoes trapped (MEL_Prop).

All models included training (.75) and a testing (.25) dataset that were
generated using the random seed 123. All analysis can be found in the
analysis.qmd file. The first model was a GAM with all eight predictors,
followed by a Random Forest, LASSO, Generalized Additive Model with
Sparsity (GAMS), and a Gradient Boosting Machine (GBM) model.

As expected, the RSME is lower for the training dataset than the testing
dataset for all models. The RMSE values are displayed in the table
below.

![](images/rmse_table-03.png){fig-align="center"}

# Discussion

The Root Mean Square Error (RMSE) comparison outlined in the table
provides a clear perspective on the efficacy of different predictive
models in the context of Eastern Equine Encephalitis (EEE) surveillance
in southeastern Massachusetts. The null model, which served as a
baseline with an RMSE of 2.40, indicates the level of error absent any
predictive influence. This baseline was pivotal for evaluating
enhancements brought by more complex models.

The Generalized Additive Model (GAM), which incorporated eight predictors selected based on their statistical significance from preliminary analyses, demonstrated concerning signs of overfitting. While the model achieved a low RMSE of 1.7 on the training data, indicating a good fit to this dataset, its performance on the testing data was substantially worse, with an RMSE of 3.44. This significant increase in RMSE when applied to the testing data, compared to the training data, suggests that the model, despite fitting the training data well, fails to generalize to new, unseen data. This discrepancy in performance is further highlighted by the testing RMSE being higher than the RMSE of the null model at 2.40, underscoring the limitations of the GAM in this scenario.

Subsequent models such as the Random Forest, LASSO, Generalized Additive
Model with Sparsity (GAMS), and Gradient Boosting Machine (GBM) were
tested to potentially mitigate overfitting and enhance prediction
accuracy. Notably, the Gradient Boosting Machine model achieved the
lowest testing RMSE of 2.14, suggesting a model with better
generalizability compared to others.

Given that the average MIR value is 1.6, an RMSE of over 2 suggests that the models are not accurately capturing the variability in the data and are making predictions that are significantly different from the observed values.

Each model's performance underlines the complex interplay of various
ecological and environmental factors in predicting EEE transmission
cycles.

## Strengths and Limitations

The accuracy and reliability of the mosquito collection data is
excellent. The same procedures and sentinel sites have were in place for
the entire analytical horizon. Further, there were only two people
responsible for sorting, identifying and counting species over the
entire period. However, the data is not without limitations. Mosquito
surveillance is inherently biased. Traps are set in locations that are
most likely to capture infected mosquitoes. While this is a positive for
early warning of virus activity, it is not necessarily representative of
the surrounding area. Probably the biggest limitation is the small
number of observations. The mosquito season is short in Massachusetts
and funding for trapping and testing mosquitoes is limited. Mosquito
control efforts, like pesticide applications, are missing from this
analysis. These efforts are designed to lower the levels of virus
transmitting mosquitoes and would likely provide more insight into the
relationship between the predictors and virus activity.

MIR is not a perfect measure of virus activity in mosquitoes. It tends
to underestimate the level of virus circulating in the environment.
However, other measures require equal pool numbers for testing. There is
a correction factor for variable pool numbers and I will explore this
option in later phases of this project. Ideally, the outcome would be
human or animal cases. However, these cases are rare and modeling them
given the limited amount of surveillance data may not be possible.

Initially, the lack of significance for most temperature and
precipitation variable was surprising. However, upon closer examination,
it makes sense. The main enzootic vector of EEE and the one targeted for
surveillance, *Culiseta melanura*, develops in well insulated crypts
that are less sensitive to ambient air temperature changes. The effect
of temperature changes on MEL development speed may be less pronounced
than other vector species that develop in open, standing water, like
ditches and containers. Precipitation increases the amount of standing
water and increases the chances of larval survival during the mosquito
season and also during the winter months for those species that
overwinter in the larval stage. For EEE, the main enzootic vector, MEL,
overwinters as larvae.

The spatial scale is an issue with the human and veterinary cases data.
Human case data is released only on a county scale to ensure patient
privacy. Some veterinary case data is available on the town-level, but
most remains only on the county-level. These limitations are barriers to
accurate spatial analysis. The project will rely on virus isolation in
mosquitoes to model risk. The available bird data relies heavily on
crowd-sourced reports. There is one standardized survey (Breeding Bird
Survey) that seems to be more accurate and reliable. However, this data
is not available for every year.

## Conclusions

While all models
demonstrated improvement over null model, none of them were robust
enough to be used for reliable prediction. In the next phases of this
project, the data will be expanded to include more years and cover a
larger portion of Massachusetts. In additon to increasing the the number
of observations, other sources of data will be incorporated. Including
virus activity from surrounding regions and incorporating bird population data in to the models could provide more insight into the relationship between the predictors and virus activity.

# References